### 2019-09-10 caleidoscope/caleidoscope.tex
done section 3

\documentclass[a4paper]{article}

%\usepackage{aiml18macro}

\usepackage[margin=2cm]{geometry}
\usepackage[noxy]{virginialake}                        
\usepackage{lmodern}
\usepackage{amssymb,amsmath,amsthm} 
\usepackage{mathtools}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{xcolor}
\usepackage{pgf}
\usepackage[all]{xy}
\usepackage{tikz}
%\usepackage{hyperref}
\usepackage{pgffor}

	%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lem}[theorem]{Lemma}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
	%%
\theoremstyle{definition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{construction}[theorem]{Construction}
\newtheorem{remark}[theorem]{Remark}
	%%

\tikzset{
	annotated cuboid/.pic={
		\tikzset{%
			every edge quotes/.append style={midway, auto},
			/cuboid/.cd,
			#1
		}
		\draw [every edge/.append style={pic actions, opacity=.5}, pic actions]
		(0,0,0) coordinate (o) -- ++(-\cubescale*\cubex,0,0) coordinate (a) -- ++(0,-\cubescale*\cubey,0) coordinate (b) edge coordinate [pos=1] (g) ++(0,0,-\cubescale*\cubez)  -- ++(\cubescale*\cubex,0,0) coordinate (c) -- cycle
		(o) -- ++(0,0,-\cubescale*\cubez) coordinate (d) -- ++(0,-\cubescale*\cubey,0) coordinate (e) edge (g) -- (c) -- cycle
		(o) -- (a) -- ++(0,0,-\cubescale*\cubez) coordinate (f) edge (g) -- (d) -- cycle;

		;
	},
	/cuboid/.search also={/tikz},
	/cuboid/.cd,
	width/.store in=\cubex,
	height/.store in=\cubey,
	depth/.store in=\cubez,
	units/.store in=\cubeunits,
	scale/.store in=\cubescale,
	width=10,
	height=10,
	depth=10,
	units=cm,
	scale=.1,
}
%tikz parameters

\tikzstyle{point}=[circle,draw]
\usetikzlibrary{arrows,automata,shapes,decorations.markings,
	decorations.pathmorphing,backgrounds,fit,snakes,calc}

%\usepackage{float}
%\floatstyle{boxed} 
%\restylefloat{figure}

\usepackage{colonequals}
\usepackage{bm}

%%% Extracting symbols from MnSymbol 
\DeclareFontFamily{U} {MnSymbolC}{}
%
\DeclareFontShape{U}{MnSymbolC}{m}{n}{
	<-6>  MnSymbolC5
	<6-7>  MnSymbolC6
	<7-8>  MnSymbolC7
	<8-9>  MnSymbolC8
	<9-10> MnSymbolC9
	<10-12> MnSymbolC10
	<12->   MnSymbolC12}{}
\DeclareFontShape{U}{MnSymbolC}{b}{n}{
	<-6>  MnSymbolC-Bold5
	<6-7>  MnSymbolC-Bold6
	<7-8>  MnSymbolC-Bold7
	<8-9>  MnSymbolC-Bold8
	<9-10> MnSymbolC-Bold9
	<10-12> MnSymbolC-Bold10
	<12->   MnSymbolC-Bold12}{}
%
\DeclareSymbolFont{MnSyC}         {U}  {MnSymbolC}{m}{n}
%
\DeclareMathSymbol{\diamondplus}{\mathbin}{MnSyC}{124}
\DeclareMathSymbol{\boxtimes}{\mathbin}{MnSyC}{117}

%%% Comments  
\definecolor{notgreen}{rgb}{.1,.6,.1}
\newcommand{\marianela}[1]{{\color{purple}[Marianela: #1]}}
\newcommand{\sonia}[1]{{\color{blue}[Sonia: #1]}}
\newcommand{\lutz}[1]{{\color{notgreen}[Lutz: #1]}}
\newcommand{\todo}[1]{{\color{red}[TODO: #1]}}

%%% General
\input{macros}
%\newcommand{\B}{\Digamma}
\newcommand{\B}{\mathcal{R}}
%\newcommand{\B}{\Xi}
%\newcommand{\B}{\mathcal{B}}
\newcommand*{\A}{\mathcal{A}}

\newcommand{\lseq}[3]{#1 , #2 \SEQ #3}

%%% Systems
\newcommand*{\ax}[1]{\mathsf{#1}}
\newcommand*{\kax}[1][]		{\ax{k_{#1}}}
\newcommand*{\lab}{\mathsf{lab}}
\newcommand*{\IK}{\mathsf{IK}}
\newcommand*{\K}{\mathsf{K}}
\newcommand*{\labIKp}{\lab\IK_{\le}}
\newcommand*{\ISfour}{\mathsf{IS4}}
\newcommand*{\Sfour}{\mathsf{S4}}

\newcommand{\fourax}{\mathsf{4}}
%%% Connectives
\newcommand*{\NOT}{\neg}
\newcommand*{\AND}{\mathbin{\wedge}}
\newcommand*{\TOP}{\mathord{\top}}
\newcommand*{\OR}{\mathbin{\vee}}
\newcommand*{\BOT}{\mathord{\bot}}
\newcommand*{\IMP}{\mathbin{\scalebox{.9}{\raise.2ex\hbox{$\supset$}}}}

\newcommand*{\BOX}{\mathord{\Box}}
\newcommand*{\DIA}{\mathord{\Diamond}}

%%% Labelled sequents
\newcommand*{\fm}[1]{{\color{notgreen}{#1}}}
\newcommand*{\lb}[1]{{\color{blue}{#1}}}
\newcommand*{\labels}[2]{\lb{#1}\mathord{:}\fm{#2}}
\newcommand*{\accs}[2]{\lb{#1}R\lb{#2}}
\newcommand*{\futs}[2]{\lb{#1}\le{\color{blue}{#2}}}
\newcommand{\SEQ}{\Longrightarrow}
\newcommand*{\DD}{\mathcal{D}}
\newcommand*{\rn}[1]  {\ensuremath{\mathsf{#1}}}
\newcommand*{\invr}[1]{#1^\bullet}
\newcommand*{\rel}{R}

\newcommand{\proviso}[1]{\mbox{\scriptsize #1}}
%%% Labelled rules
\newcommand*{\labrn}[2][]  {\rn{#2}_{#1}}%^{\lab}}}
\newcommand*{\rlabrn}[2][]  {\rn{#2}_\rn{R#1}}%^\lab}}
\newcommand*{\llabrn}[2][]  {\rn{#2}_\rn{L#1}}%^\lab}}
%
\newcommand*{\brsym}{\boxtimes}%\mathord{\scalebox{.8}{$\blacksquare$}}}
\newcommand*{\diasym}{\diamondplus}%\mathord{\blacklozenge}}
%
\newcommand*{\boxbrn}[1]{\brsym_\rn{#1}}%^{\lab}}}
\newcommand*{\diabrn}[1]{\diasym_\rn{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Equation environments
\newdimen\mydisplayskip
\mydisplayskip=.4\abovedisplayskip
\newenvironment{smallequation}
{\par\nobreak\vskip\mydisplayskip\noindent\bgroup\small\csname equation\endcsname}{\csname endequation\endcsname\egroup}
\newenvironment{smallequation*}
{\par\nobreak\vskip\mydisplayskip\noindent\bgroup\small\csname equation*\endcsname}{\csname endequation*\endcsname\egroup}
\newenvironment{smallalign}
{\par\nobreak\noindent\bgroup\small\csname align\endcsname}{\csname endalign\endcsname\egroup}
\newenvironment{smallalign*}
{\par\nobreak\noindent\bgroup\small\csname align*\endcsname}{\csname endalign*\endcsname\egroup}
%
\newcommand{\vlderivationauxnc}[1]{#1{\box\derboxone}\vlderivationterm}
\newcommand{\vlderivationnc}{\vlderivationinit\vlderivationauxnc}
%
%
\makeatletter
\newbox\@conclbox
\newdimen\@conclheight
%
%%%%%%
\newcommand{\vlhtr}[2]{\vlpd{#1}{}{#2}}
\newcommand\vlderiibase[5]{{%
		\setbox\@conclbox=\hbox{$#3$}\relax%
		\@conclheight=\ht\@conclbox%
		\setbox\@conclbox=\hbox{$%
			\vlderivationnc{%
				\vliin{#1}{#2}{\box\@conclbox}{#4}{#5}%
			}$}%
		\lower\@conclheight\box\@conclbox%
	}}
%%
\newcommand\vlderibase[4]{{%
		\setbox\@conclbox=\hbox{$#3$}\relax%
		\@conclheight=\ht\@conclbox%
		\setbox\@conclbox=\hbox{$%
			\vlderivationnc{%
				\vlin{#1}{#2}{\box\@conclbox}{#4}%
			}$}%
		\lower\@conclheight\box\@conclbox%
	}}
%
\newcommand\vlderidbase[4]{{%
		\setbox\@conclbox=\hbox{$#3$}\relax%
		\@conclheight=\ht\@conclbox%
		\setbox\@conclbox=\hbox{$%
			\vlderivationnc{%
				\vlid{#1}{#2}{\box\@conclbox}{#4}%
			}$}%
		\lower\@conclheight\box\@conclbox%
	}}
	%
	\makeatother
	%
%%%General maths
\newcommand*{\reducesto}{\quad{\leadsto}\quad}

%%%%%%%%%%%%%%%%%

\newcommand{\qquand}{\qquad\mbox{and}\qquad}
\newcommand{\quadcm}{\rlap{\quad,}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%The following line defines the page header consisting of the surnames of the authors.
% Please include only the last names! 
% Separate by commas except the last two surnames which are separated by an "and".
\def\lastname{Marin, Morales and Stra{\ss}burger}

\begin{document}

%\begin{frontmatter}
  \title{A properly labelled proof system for intuitionistic modal logics}
  \author{Sonia Marin \and %}
%  \todo{New university?}
%  \address{IT-Universitetet i K{\o}benhavn \\ Denmark }
%    \author{
      Marianela Morales %}
 \and
% \marianela{I think that now I have the same as Lutz, right?} No, the other order...
% \address{ LIX, \'Ecole Polytechnique  \& Inria Saclay\\ France}
%      \author{
        Lutz Stra{\ss}burger }
% \address{Inria Saclay \& LIX, \'Ecole Polytechnique \\ France}

  \maketitle

 \begin{abstract}

   In this paper we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols for the labels: one for the accessibility
   relation associated with the Kripke semantics for modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logic. As a
   consequence it can encompass a wider range of intuitionistic modal
   logics than standard labelled systems.
   %%
   We also show an internal cut elimination proof for our system.

   %% . We show the possibility of extend labelled sequents
   %% with a preorder relation symbol in order to capture intuitionistic
   %% modal logic. For this, we obtain a proof system which is complete
   %% with respect to Hilbert system.  Also, we present the proof for
   %% completeness using Simpson system.
  \end{abstract}

  %% \begin{keyword}
  %% Intuitionistic modal logic, labelled sequents, cut elimination, proof theory.
  %% \end{keyword}
% \end{frontmatter}

  %\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{get bib-file right}

\section{Introduction}\label{sec:intro}

Since their introduction in the 1980s by Gabbay~\cite{Gabbay},
\emph{labelled proof calculi} have been widely used by proof theorist
to give sound, complete and cut-free deductive system to a broad range
of logics. Unlike so-called \emph{internal calculi}, like
hyper-sequents~\cite{avron} or nested
sequents~\cite{kashima,brunnler,pog} or 2-sequents~\cite{masini},
labelled calculi have the advantage of beeing more uniform and
accommodating more logics.

Standard labelled sequent calculi attach to every formula $A$ a label
$x$, witten as $\labels xA$, and additionally admit so called
\emph{relational atoms} of the form $\accs xy$ where $\rel$ is a
binary relation symbol. These calculi work best for logics with
standard Kripke semantics, as in this case the relation $\rel$ is used
to encode the accessibility relation in the Kripke models, and the
frame condition corresponding to the desired logic can be directly
encoded as inference rules. Prominent examples are classical modal
logics and intuitionistic propositional logic, where, e.g., the frame
condition of transitivity ($\forall \lb {xyz}.\:\accs xy\AND \accs yz\IMP\accs xz$), can be straightforwardly translated into the
inference rule
\begin{equation}
  \label{eq:Rtrans}
  \vlinf{\rn{trans}}{}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
    \B, \accs xy, \accs yz, \accs xz, \Left \SEQ \Right}
\end{equation}
where $\B$ stands for a set of relation atoms, and $\Gamma$ and
$\Delta$ for sets \lutz{or multisets?} of labelled formulas.

However,
in this paper we are concerned with \emph{intuitionistic modal
  logics}, whose Kripke semantics is based on \emph{birelational}
frames, i.e., they have two binary relations instead of one: one
relation $R$ that corresponds to the accessibility relation in
ordinary Kripke frames for modal logics, and a relation $\le$ that
corresponds to the future relation in ordinary Kripke frames for
intuitionistic logic. Consequently, standard labelled systems for
these logics have certain shortcomings:
\begin{enumerate}
\item The transitivity rule in~\eqref{eq:Rtrans} corresponds to both
  versions of the $\fourax$-axiom
  \begin{equation}
    \label{eq:4ax}
    \fm{\BOX A\IMP \BOX\BOX A}
    \qquand
    \fm{\DIA\DIA A\IMP\DIA A}
    \quadcm
  \end{equation}
  as they are equivalent in classical
  modal logic. However, in intuiotionistic modal logic they are not
  equivalent, and even though the logic $\ISfour$, i.e., the
  intuitionistic version of the modal logic $\Sfour$, contains both
  axioms, they can be added independently to the logic $\IK$ (the
  intuitionistic version of $\K$). However, the proof theory of these
  new logics has not been studied before, as no existing labelled (or
  label-free) proof system can handle them, even though the
  corresponding frame conditions
  \begin{equation}
    \label{eq:4frame}
    \forall\lb{xyz}.\:\accs xy\AND \accs yz\IMP(\exists\lb{x'}.\:\futs x{x'}\AND\accs {x'}z)
    \qquand
    \forall\lb{xyz}.\:\accs xy\AND \accs yz\IMP(\exists\lb{z'}.\:\futs z{z'}\AND\accs {x}{z'})
  \end{equation}
  have already been studied
  in~\cite{Plotkin}.\todo{check: did they get that wrong in  \cite{Plotkin}?}
\item The correspondence between the syntax and the semantics is not
  as clean as one would expect. As only the $R$-relation (and not the
  $\le$-relation) of the frame is visible in an ordinary labelled
  sequent, we only have that a sequent $\Gamma$ is provable if and
  only if is is valid in all \emph{graph-consistent}\footnote{This
    means that every layer in the model can be lifted to any future of
    any world in that layer.}\todo{check terminology in Simpson} models, as
  already observed in by Simpson in his PhD thesis~\cite{simpson:phd}
  and considered as a very unsatisfactory situation\todo{check what he
    said exactly} (see also \cite{mar:str:indexed}).
\end{enumerate}
In order to address these two issues we propose here to extend the
notion of labelled sequent by also allowing relational atoms of the
form $\futs xy$. Consequently, we can easily translate the frame
conditions in~\eqref{eq:4frame} into inference rules:
\begin{equation}
  \label{eq:4rules}
  \vlinf{\rn{4^{\BOX}}}{\proviso{$\lb{x'}$ fresh}}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
     \B, \accs xy, \accs yz, \accs {x'}z, \futs x{x'},\Left \SEQ \Right}
  \qquand
  \vlinf{\rn{4^{\DIA}}}{\proviso{$\lb{z'}$ fresh}}{\B, \accs xy, \accs yz, \Left \SEQ \Right}{
    \B, \accs xy, \accs yz, \accs {x}{z'}, \futs z{z'},\Left \SEQ \Right}  
\end{equation}
This allows us to define cut-free deductive systems for a wide range
of logics that could not be treated before.  Furthermore, the relation
between syntax and semantics is as one would expect: A sequent is
provable in our system if and only if it is valid in all models.

Besides that, there is another pleasant observation to make about our
system. Ordinary labelled sequent systems for intuitionistic modal
logic are single-conclusion~\cite{simpson:phd}. The same is true for
the corresponding nested sequent
systems~\cite{str:fossacs13,mar:str:aiml}. It is possible to express
Maehara style multiple-conclusion systems in nested
sequents~\cite{kuz:str:maehara}, and therefore also in ordinary
labelled sequents. However, also in these systems there are rules
($\rlabrn\IMP$ and $\rlabrn\BOX$) that force a single-conclusion
premise, even though this is not the case in labelled
systems~\cite{negri:xx} or nested sequents~\cite{fitting:xx} for
intuitionistic logic. Our system eliminates this
discrepancy. Consequently, every rule in our system is invertible, as
we never delete information.\lutz{check, is this really the case?}

\medskip

This paper is organized as follows, In the next section (Section~\ref{sec:intmod}) we recall the standard syntax and semantics of intuitionistic modal logics. Then, in Section~\ref{sec:system} we present our system for the intuitionistic modal logic $\IK$. In Sections~\ref{sec:soundness} and~\ref{sec:completeness}, we show soundness and completeness of the system with cut. The cut elimination theorem, proved in Section~\ref{sec:cutelim}, then entails soundness and completeness for the cut-free system. Finally, in Section~\ref{sec:ext} we discuss the possible extension to the system to capture other intuitionistic modal logics. \lutz{what about doing that directly in the system presentation, like Negri does it?}

\todo{maybe paragraph on related work? cite Maffezioli, Naibo and Negri~\cite{Maffezioli}.}

\section*{Introduction(old version, to be deleted)}
\todo{We should work on the introduction}

One possible-world semantics was established as a solid base to define modal logics, the idea of incorporating these notions into the proof theory of modal logics emerged. Fitch seems to have been the first one to formalise it, directly including symbols representing worlds into the language of his proofs in natural deduction \cite{Fitch}.

\emph{Labelled deduction} has been proposed by Gabbay~\cite{Gabbay} in the 80s as a unifying framework throughout proof theory in order to provide proof
systems for a wide range of logics. 
%
For modal logics it can take
the form of labelled natural deduction and labelled sequent systems as
used, for example, by Simpson~\cite{Simpson}, Vigan\`o~\cite{Vigano}, and
Negri~\cite{Negri}. 
%
These formalisms make explicit use not only of
labels, but also of relational atoms referring to the accessibility relation of a Kripke frame.
%
In this short note we propose a system that represents both the \emph{accessibility relation} (for modal
logics) and the \emph{preorder relation} (for intuitionistic
logic), using the full power of the bi-relational semantics for
intuitionistic modal logics,
and developing fully the idea that Maffezioli, Naibo and Negri employed in~\cite{Maffezioli}.

\todo{Simpson does not have internal cut-elimination}

\sonia{
  \begin{itemize}
  \item internal vs externl calculi
  \item extesion to axioms are truly intuitionistic (not just the
    classical versions as in Simpson style)
  \item connection with semantics compared to graph homomorphism
    restiction in Simpson and ``indexed nested systems''
  \end{itemize}
}

\lutz{things that should be said explicitely somewhere (maybe intro, maybe somewhere else):
  \begin{itemize}
  \item what we do:
    \begin{itemize}
    \item first systems (nested and labelled) for intuitionistic modal
      logic where ``single conclusion'', i.e., only one formula on the
      right.
    \item Then ``multiple conclusion'' (Meahara-style) nested sequent.
    \item Still rules that forced single formula on the right (white implication and white box)
    \item by allowing both relations of the model in the proof system
      we go to genuinely multiple conclusion: never ever a formula on
      the right has to be deleted during proof search.
    \end{itemize}
  \item why we do it: 
    \begin{itemize}
    \item getting things right: labelled systems for classical modal
      logics, or plain propositional intuiionistic logic have one
      binary relation, which is exactly the relation of the standard
      Kripke semantics. The standard Kripke semantics for
      intuitionistic modal logic is birelational, with a strong
      connectiion between the two binary relations. A clean labelled
      system for these logics should incorporate both relations.
    \item We get new logics.  In the standard nested sequent
      or labelled sequent system for intuitionistic modal logic we
      always have both conjuncts we we add Scott-Lemmon/Geach axioms;
      but here we can have them separtely.
    \end{itemize}
  \item how we do it:
    \begin{itemize}
    \item We start from the Maehara nested sequent system (which can
      be translated straightforwardly into a labelled system with one
      relation $R$.
    \item the rules for $\AND$ and $\OR$ are trivial and as expected
      they can be presented such that they are invertible on the left
      and the right, they do not concern the relations $R$ and $\le$. 
    \item So should be the rules for $\IMP$-left and $\BOX$-left and
      identity, but there is a subtlety; see below...
    \item the rules for $\DIA$ are also straightforward, as they
      concern only $R$ but not $\le$.
    \item The rules $\BOX$-right and $\IMP$-right are the ones that
      force single conclusion in the Maehara system. This means that
      here they create a future, using the relation $\le$.
    \item In order to make everything work together, this forces us to
      have reflexivity and transitivity of $\le$, explicely as rules
      in the system, and have rules for the F1 and F2 conditions
      making the relations $R$ and $\le$ go together.
    \item As a consequence we also need a rule realizing monotonicity
      for the formulas on the right.
    \item However, we designed the system such that monotonicity is
      admisible by encorporating it into the rules for identity,
      $\IMP$-left and $\BOX$-left.
    \end{itemize}
  \end{itemize}
}
%
%For sequent systems for intuitionistic logics there is always a choice
%to be made: make the system \emph{single conclusion} following
%Gentzen~\cite{Gentzen} or \emph{multiple conclusion} following
%Maehara~\cite{Maehara}. In our work we choose the multiple conclusion
%variant because of the closer correspondence to the semantics. In that
%respect, our system is closer to~\cite{kuz:str} than
%to~\cite{Simpson} and~\cite{mar:str}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section{Preliminaries}
%\marianela{Depends on how long the paper has to be, we can add a section with preliminaries: labelled sequents for classical modal logic and syntax and semantics of classical modal logic}

%\todo{re-build this section since the section \emph{intuitionistic modal logics} is part of the preliminaries}

%\sonia{We should not have the definitions of classical modal logic. Just the intuitionistic part. If we want to say something else about labelled sequents, we should add it before the definition of \emph{a two-sided labelled sequent}}

%Labelled sequents are formed from by labelled formulas of the form $\labels{x}{A}$  and relational atoms of the form $xRy$, where $x$, $y$ range over a set of variables and $A$ is a modal formula. A one-sided labelled sequent is then of the form $\B \SEQ \Right$ where $\B$ denotes a set of relational atoms and $\Right$ a multiset of labelled formulas. A simple proof system for classical modal logic K can be obtained in this formalism (\ref{fig:systemlabK}).

%%\begin{figure}[h]
%\begin{center}

%$\vlderivation {\vlinf{\id}{}{\B \SEQ \Right, x: a, x: \vls-a}{}}$
%\hspace{7mm}$\vlderivation {\vlinf{\tolab}{}{\toprule}{}}$

%$\vliinf{\vlab}{}{\B \SEQ \Right, x :\vls(A.B)}{\vlabr}{\vlabu}$
%\hspace{7mm}$\vlinf{\olab}{}{\B \SEQ \Right, x \colon \vls[A.B]}{\olabr}$

%$\vlinf{\blab}{y$ fresh$}{\blabr}{\blabu}$
%\hspace{7mm}$\vlinf{\dlab}{}{\dlabr}{\dlabu}$

%\end{center}
%\caption{System $\mathsf{labK}$}
%\label{fig:systemlabK}
%\end{figure}

%These rule schemes can occur in different contexts and different calculi. The context that interests us is when it is applied to modal logic. The idea to extend this system to another to capture intuitionistic modal logics allows us to study the Kripke semantics for this type of logics:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Intuitionistic modal logics}\label{sec:intmod}

	The language of {intuitionisitic modal logic} is the one of intuitionistic propositional logic with the modal operators $\BOX$ and $\DIA$, standing most generally for \emph{necessity} and \emph{possibility}.
	%
	Starting with a set $\mathcal{A}$ of atomic propositions, denoted $a$, modal formulas are constructed from the grammar:
	%
	$$
	A \coloncolonequals
	a \mid A \AND A \mid \TOP \mid A \OR A \mid \BOT \mid A \IMP A \mid \BOX A \mid \DIA A
	$$
	%
        \lutz{why do we have $\TOP$?}
	We might sometimes write $\NOT A$ to mean $A \IMP \BOT$. \lutz{do we?}

Obtaining the intuitionistic variant of $\mathsf{K}$ is more involved than the classical variant. 
%%
Lacking De Morgan duality, there are several variants of $\mathsf{k}$ that are classically but not intuitionistically equivalent. Five axioms have been considered as primitives in the literature. 
%

The axiomatisation that is now generally accepted as intuitionistic modal logic $\IK$ was given by Plotkin and Stirling~\cite{Plotkin} and is equivalent to the one proposed by Fischer-Servi~\cite{Fischer}.
%, and by Ewald~\cite{Ewald} in the case of intuitionistic tense logic. 
%
It is obtained from intuitionistic propositional logic by adding:
\begin{itemize}
\item the \emph{necessitation rule}: if $A$ is a theorem then $\BOX A$ is also a theorem; and
\item the following five variants of the \emph{distributivity axiom}:
\begin{equation*}
\label{eq:ik}%\hskip-2em
\begin{array}[t]{r@{\;}l@{\quad}r@{\;}l@{\quad}r@{\;}l}
\kax[1]\colon&\BOX(A\IMP B)\IMP(\BOX A\IMP\BOX B)
&
\kax[3]\colon&\DIA(A\OR B)\IMP(\DIA A\OR\DIA B)
&
\kax[5]\colon&\DIA\BOT\IMP\BOT
\\
\kax[2]\colon&\BOX(A\IMP B)\IMP(\DIA A\IMP\DIA B)
&
\kax[4]\colon&(\DIA A\IMP \BOX B)\IMP\BOX(A\IMP B)\\%x[1ex]
\end{array}
\end{equation*}
\end{itemize}

The relational semantics for $\IK$ was first defined by Fischer-Servi~\cite{Fischer}.
%
It combines the Kripke semantics for intuitionistic propositional logic and the one for classical modal logic, using two distinct relations on the set of worlds.

\begin{definition}
	A \emph{bi-relational frame} $\F$ is a triple $\langle W, R, \le \rangle$ 
%	of a non-empty set of worlds $W$ equipped with two binary relations $R$ and $\le$, where $R$ being the modal \emph{accessibility relation} and $\le$ a preorder (\emph{i.e.} a reflexive and transitive relation), satisfying the following conditions:
	of a set of worlds $W$ equipped with an {accessibility relation} $\rel$ and a preorder $\le$ (\emph{i.e.} a reflexive and transitive relation) satisfying:
	\begin{enumerate}
		\item[($F_1$)] For $u, v, v' \in W$, if $u \rel v$ and $v \le v'$, there exists $u'$ s.t.~$u \le u'$ and $u' \rel v'$.

			\begin{equation*}
			\xymatrix{
				u' \ar@{.>}[r]_R  & v' \\
				u \ar@{.>}[u]^{\le} \ar@{->}[r]_R  & v \ar@{->}[u]_{\le}
			}
			\end{equation*}

		\item[($F_2$)] For $u', u, v \in W$, if $u \le v$, there exists $v'$ s.t.~$u' \rel v'$ and $v\le v'$.

			\begin{equation*}
			\xymatrix{
				u' \ar@{.>}[r]^R & v' \\
				u \ar@{->}[u]^{\le} \ar@{->}[r]^R & v \ar@{.>}[u]_{\le}
			}
			\end{equation*}
	\end{enumerate}
%	
\end{definition}

\begin{definition}
	\label{model}
	A \emph{bi-relational model} $\M$ is a quadruple $\langle W, R,\le,V \rangle$ with $\langle W, R, \le \rangle$ a bi-relational frame and $V\colon W \to 2^\mathcal{A}$ a monotone valuation function, that is, a function mapping each world $w$ to the subset of propositional atoms true at $w$, additionally subject to:
	if $w \le w'$ then $V(w)\subseteq V(w')$.
\end{definition}

We write $w \Vdash a$ if $a \in V(w)$, and by definition, we always have $w \Vdash \top$ and never that $w \Vdash \bot$. 
%
Then the relation is extended to all formulas by induction, following the rules for both intuitionistic and modal Kripke models:
\begin{equation}
  \label{eq:kripke}
  \parbox{30em}{%
  $w \Vdash A \AND B$ iff $w \Vdash A$ and $w \Vdash B$

$w \Vdash A \OR B$ iff $w \Vdash A$ or $w \Vdash B$

$w \Vdash A \IMP B$ iff for all $w'$ with $w \le w'$, if $w' \Vdash A$ then $w' \Vdash B$

$w \Vdash \BOX A$ iff for all $w'$ and $u$ with $w \le w'$ and $w'Ru$, $u \Vdash A$ \hfill %$(\ast)$

$w \Vdash \DIA A$ iff there exists a $u$ such that $wRu$ and $u \Vdash A$.

%We write $w \not\Vdash A$  if it is not the case that $w\Vdash A$.\\
}
\end{equation}
\todo{fix typography}

We write $w \not\Vdash A$ if it is not the case that $w \Vdash A$, but contrarily to the classical case, we do not have $w \Vdash \neg A$ iff $w \Vdash A$ (since $\neg A$ is defined as $A \IMP \bot$).

From the monotonicity of the valuation function $V$, we get a monotonicity property for the relation:
\begin{proposition}(Monotonicity) 
For any formula $A$ and for $w, w' \in W$, if $w \le w'$ and $w \Vdash A$, then $w' \Vdash A$.
\end{proposition}

\begin{definition}
	A formula $A$ is \emph{satisfied} in a model $\M = \langle W, R, \le, V \rangle$, if for all $w \in W$ we have $w \Vdash A$.
%
	A formula $A$ is \emph{valid} in a frame $\F = \langle W, R, \le \rangle$, if for all valuations $V$, $A$ is satisfied in $\langle W, R, \le, V \rangle$.
\end{definition}

%It then was investigated in detail in~\cite{Simpson}, in which strong arguments are given in favour of this axiomatic definition: 
%
%it allows for adapting to intuitionistic logic the standard embedding of modal logic into first-order logic, and also provides an extension of the standard Kripke semantics for classical modal logic to the intuitionistic case.

Similarly to the classical case, in the case of $\IK$, the correspondence between syntax and semantics is recovered.

\begin{theorem}[Fischer-Servi~\cite{Fischer}, Plotkin and Stirling~\cite{Plotkin}]\label{thm:plotkin}
	A formula $A$ is a theorem of $\IK$ if and only if $A$ is valid in every bi-relational frame.
\end{theorem}

\todo{Add something else about Fischer-Servi theorem?}\lutz{no, but we should add the Plotkin-Stirling theorem about the relation between axioms and frame conditions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The system}\label{sec:system}

In this section we present our properly labelled sequent proof system
for intuitionistic modal logics. The starting point is the notion of a
\emph{labelled formula} which is a pair $\labels xA$ of a label $\lb
x$ and a formula $\fm A$. The intended meaning is that $\fm A$ holds
at world $\lb x$. A \emph{relation atom} is either an expression
$\accs xy$ or $\futs xy$ where $\lb x$ and $\lb y$ are labels and the
intended meaning is that the corresponing relation holds between the
worlds $\lb x$ and $\lb y$. This is enough to define the notion of
\emph{(labelled) sequent}, which is a triple $\lseq\B\Left\Right$,
where $\B$ is a set of relation atoms, and $\Left$ and $\Right$ are
sets of labelled formulas, all written as list, separated by commas.

\begin{figure}[!t]
  \begin{center}
	\fbox{
%	\begin{minipage}{.95\textwidth}
	\begin{tabular}{c@{\quad}c}
		$\vlinf{\rn{id}}{}{\B, x \le y, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a} }{}$
		&
		$\vlinf{\llabrn\bot}{}{\B, \Left, \labels{x}{\BOT} \SEQ \Right}{}$
		\quad
		$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, \labels{x}{\TOP}}{}$
		\\\\
		$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
		&
		$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, \labels{x}{A \AND B}}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{B}}$
		\\\\
		$\vliinf{\llabrn\OR}{}{\B, \Left, \labels{x}{A \OR B} \SEQ \Right}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{B} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rlabrn\IMP}{\text{\scriptsize $y$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{\B, x \le y, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$
		}
		\\\\
		\multicolumn{2}{c}{
 		$\vliinf{\llabrn\IMP}{}{\B, x \le y, \Left, \labels{x}{A} \IMP B \SEQ \Right}{\B, x \le y, \labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{y}{A}}{\B, x \le y, \Left, \labels{y}{B} \SEQ \Right}$ 
% 		&$\vliinf{\rlabrn\IMP}{\text{\scriptsize $x \le y \in \B$}}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{\B, \Left \SEQ \Right, \labels{y}{A}}{\B, \Left, \labels{y}{B} \SEQ \Right}$
 		}
		\\\\
		$\vlinf{\llabrn\BOX}{}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\BOX}{\text{\scriptsize $y, z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, x \le y, y \rel z, \Left \SEQ \Right, \labels{z}{A}}$
		\\\\
		$\vlinf{\llabrn\DIA}{\text{\scriptsize $y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}$
		&
		$\vlinf{\rlabrn\DIA}{}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
		\\
		\multicolumn{2}{c}{
		$\mbox{\hbox to .9\linewidth{\dotfill}}$
		}
		\\
		$\vlinf{\rn{refl}}{}{\B, \Left \SEQ \Right}{\B, x\le x, \Left \SEQ \Right}$
		&
		$\vlinf{\rn{trans}}{}{\B, x \le y, y \le z, \Left \SEQ \Right}{\B, x \le y, y \le z, x \le z, \Left \SEQ \Right}$
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rn{F_1}}{\text{\scriptsize $u$ fresh}}{\B, xRy, y \le z, \Left \SEQ \Right}{\B, xRy, y \le z, x \le u, uRz, \Left \SEQ \Right}$
		}
		\\\\
		\multicolumn{2}{c}{
		$\vlinf{\rn{F_2}}{\text{\scriptsize $u$ fresh}}{\B, xRy,x \le z, \Left \SEQ \Right}{\B, xRy, x \le z, y \le u, zRu, \Left \SEQ \Right }$		
		}
	\end{tabular}		
%\end{minipage}
}		
  \end{center}
  \caption{System $\labIKp$}
	\label{fig:labIKp}
\end{figure}

Now we can present the inference rules for \emph{system $\labIKp$} for the logic $\IK$.
We obtained this system, shown in Figure~\ref{fig:labIKp}, as follows.
Our starting point was the multiple-conlusion nested sequent system \emph{\`a la} Maehara (as presented in~\cite{kuz:str:maehara}), which can straightforwardly translated into the labelled setting, and which yields the rules $\llabrn\bot$ \todo{$\TOP$ yes/no}, $\llabrn\AND$, $\rlabrn\AND$, $\llabrn\OR$, $\rlabrn\OR$, $\llabrn\DIA$, and $\rlabrn\DIA$ as we show them in Figure~\ref{fig:labIKp}. However, this translation also yields the rules $\rn{id'}$, $\llabrn\IMP'$, and~$\llabrn\BOX'$:
\begin{equation}
  \label{eq:nomon}
  \vlinf{\rn{id}'}{}{\B, \Left, \labels{x}{a} \SEQ \Right, \labels{x}{a} }{}
  \quad
  \vliinf{\llabrn\IMP'}{}{\B,\Left, \labels{x}{A} \IMP B \SEQ \Right}{\B,\labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left, \labels{x}{B} \SEQ \Right}
  \quad
  \vlinf{\llabrn\BOX}{}{\B, xRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, xRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}
\end{equation}
that are not sufficient for a complete system, and we will see below
why. Before, let us first look at the rules $\rlabrn\IMP$ and
$\rlabrn\BOX$. In the multiple-conlusion nested sequent system
of~\cite{kuz:str:maehara}, these are the two rules that force
single-conclusion. This is the reason why in our system we need the
$\le$ in the relational atoms; we need to go to the future of the
current world.  In the Kripke-semantics in~\eqref{eq:kripke} the two
connectives $\IMP$ and $\BOX$ are the ones that need the
future-relation $\le$. This relation is reflexive and transitive. In
order to capture that in the proof system, we need to add the rules
$\rn{refl}$ and $\rn{trans}$.

Finally, in the semantics, the two relations $\rel$ and $\le$ are
strongly connected through the two conditions $F_1$ and $F_2$
\todo{proper typography}. These need to be reflected in the proof
system, which is done by the two rules $\rn{F_1}$ and $\rn{F_2}$.
%
However, these two rule create new labels, and in order to be
complete, the system needs the \emph{monotonicity} rule:
\begin{equation}
  \label{eq:mono}
  \vlinf{\rn{mon}}{}{\lseq{\B, x \le x'}{\Left, \labels{x}{A}}\Right}{
    \lseq{\B, x \le x'}{\Left, \labels{x}{A}, \labels{x'}{A}}\Right}
\end{equation}
Since this rule is a form of contraction, it would cause the same
problems as contraction in a cut elimination proof. Hence, it is
preferable to have a system in which this rule is admissible. This is
the reason why we have monotonicity incorporated in the rules
$\rn{id}$, $\llabrn\IMP$ and $\llabrn\BOX$ in Figure~\ref{fig:labIKp},
instead of using the rules in~\eqref{eq:nomon}.

\begin{theorem}
  \label{thm:mon-adm}
  The rule $\rn{mon}$ is admissible for system $\labIKp$.
\end{theorem}

For proving this theorem, we need a series of auxiliary lemmas.

\todo{}

\begin{proof}[Proof of Theorem~\ref{thm:mon-adm}]
  \todo{}
\end{proof}

In the following sections, we will show that the system $\labIKp$ is sound and complete. For the completeness proof we proceed via cut elimination. The cut rule has the following shape:
\begin{equation}
  \label{eq:cut}
  \vliiinf{\labrn{cut}}{}{
    \lseq\B\Left\Right}{
    \lseq\B\Left{\Right,\labels{z}{C}}}{}{
    \lseq\B{\Left,\labels{z}{C}}{\Right}}
\end{equation}
Then we can summarize soundness and completeness, and cut
admissibility of $\labIKp$ in the following Theorem:

\begin{theorem}\label{thm:cutfree-compl}
	%	Let $\CC$ be a set of geometric frame properties as in~\eqref{eq:cla-geometric} and $\labbrn{\CC}$ be the corresponding set of rules following schema~\eqref{eq:modal-grs}.
	%
	For any formula $A$, the following are equivalent.
	%
	\begin{enumerate}
		\item\label{i} $A$ is a theorem of $\IK$. 
		%
		\item\label{ii} $A$ is provable in $\labIKp +\labrn{cut}$.
		\item\label{iii} $A$ is provable in $\labIKp$.
		%
		\item\label{iv} $A$ is valid in every birelational frame. %satisfying the properties in $\CC$.
	\end{enumerate}
\end{theorem}

The proof of this theorem is the topic of the following sections. The
equivalence of \ref{i} and \ref{iv} has already been stated in
Theorem~\ref{thm:plotkin}~\cite{Fischer,Plotkin}. The implication
\ref{i}$\implies$\ref{ii} is shown in Section~\ref{sec:completeness},
the implication~\ref{ii}$\implies$\ref{iii} is shown in
Section~\ref{sec:cut-elim}, and finally, the implication
\ref{iii}$\implies$\ref{iv} is shown in Section~\ref{sec:soundness}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Soundness}\label{sec:soundness}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Completness}\label{sec:completeness}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Cut Admissibility}\label{sec:cut-elim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Labelled sequent calculi with $\rel$ and $\le$}

%%
%This worked as well in the labelled setting as in the ordinary sequent case: we followed Simpson's sequent system where intuitionistic labelled sequents are written $\B, \Left \SEQ \labels{z}{C}$ for some multiset of labelled formulas $\Left$, some formula $C$, some label $z$ and a set of relational atoms $\B$. 
Once possible-world semantics was established as a solid base to define modal logics, the idea of incorporating these notions into the proof theory of modal logics emerged.  Labelled deduction has been more generally proposed by Gabbay in the 80's as a unifying framework throughout proof theory in order to provide proof systems for a wide range of logics \cite{Gabbay}. For modal logics it can also take the form of labelled natural deduction and labelled sequent systems as used, for example, by Simpson \cite{Simpson}, Vigano \cite{Vigano}, and Negri\cite{Negri}. These formalisms make explicit use not only of labels, but also of relational atoms (which, to our knowledge, are not commonly used in tableaux systems, except for the pioneering work of Nerode \cite{Nerode}).

Structural proof theoretic accounts of intuitionistic modal logic can adopt the paradigm of \emph{labelled deduction} in the form of labelled natural deduction and labelled sequent systems~\cite{Simpson}, or the one of \emph{unlabelled deduction} in the form of sequent~\cite{Bierman} or nested sequent systems~\cite{Strassburger} (for a survey see~\cite[Chap.~3]{Marin}).

Simpson~\cite{Simpson} followed the lines of Gentzen in a labelled context, namely, he developed a labelled natural deduction framework for modal logics and then converted it into sequent systems with the consequent restriction to one formula on the right-hand side of each sequent. Simpson's proposed labelled sequent system for intuitionistic modal logic is representing explicitly only the accessibility relation $\rel$ in the syntax.
%

Echoing the definition of bi-relational structures, we consider another extension of labelled deduction to the intuitionistic setting. 
%
The idea is to use two sorts of relational atoms, one for the modal relation $\rel$ and another one for the intuitionistic relation $\leq$ as proposed by Maffezioli, Naibo and Negri in~\cite{Maffezioli}. 
%

\begin{definition}
A two-sided intuitionistic \emph{labelled sequent} is of the form $\B, \Left \SEQ \Right$ where $\B$ denotes a set of relational atoms $x \rel y$ and preorder atoms $x \le y$, and $\Left$ and $\Right$ are multi-sets of labelled formulas $\labels{x}{A}$ (for $x$ and $y$ taken from the set of labels and $A$ an intuitionistic modal formula).
\end{definition}
In the next section, we present the system (called $\labIKp$) that we obtained to capture intuitionistic modal logics (see Figure \ref{fig:labIKp}).

\section{Building $\labIKp$}

As we mentioned, a labelled sequent calculus for classical modal logic was introduced in \cite{Negri}. In our work, we show an extension of Negri's system in order to capture intuitionistic modal logics. This new system is formed from formulas of the form $\labels{x}{A}$ where $A$ is an intuitionistic modal formula, relational atoms $xRy$ ($R$ is the modal accesibility relation) where $x$, $y$ range over a set of variables called \emph{labels} and preorder atoms of the form $x \le y$. In this section, we study each rule of the system $\labIKp$ (Figure \ref{fig:labIKp}) in details.

\subsection{Frame conditions}
%In order to capture intuitionistic modal logics, we need to answer the requirements of the de
The semantics of intuitionistic modal logic introduces a bi-relational model with two binary relations: $R$ and $\le$, where $\le$ is a preorder relation, i.e, a reflexivity and transitivity relation. For that reason, the system $\labIKp$ contains the rules $\rn{refl}$ and $\rn{trans}$. Furthermore, our system has to incorporate the two semantic conditions into deductive rules as $\rn{F1}$ and $\rn{F2}$ . On the other hand, the valuation function $V$ that introduces the \emph{monotonicity property} is captured with the rule $\rn{id}$ in our system. All these are presented in the system $\labIKp$ as follows:

\begin{center}
		\begin{minipage}{.95\textwidth}
			\begin{tabular}{@{\!}c@{\quad}c}
				\multicolumn{2}{c}{
					\hspace{18mm}
					$\vlinf{\rn{id}}{}{\B, x \le y, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a} }{}$
				}
				\\\\
				\hspace{15mm}
				$\vlinf{\rn{refl}}{}{\B, \Left \SEQ \Right}{\B, x\le x, \Left \SEQ \Right}$
				&
				$\vlinf{\rn{trans}}{}{\B, x \le y, y \le z, \Left \SEQ \Right}{\B, x \le y, y \le z, x \le z, \Left \SEQ \Right}$

				\\\\
				\multicolumn{2}{c}{
					\hspace{18mm}
					$\vlinf{\rn{F_1}}{\text{\scriptsize $u$ fresh}}{\B, xRy, y \le z, \Left \SEQ \Right}{\B, xRy, y \le z, x \le u, uRz, \Left \SEQ \Right}$
				}
				\\\\
				\multicolumn{2}{c}{
					\hspace{18mm}
					$\vlinf{\rn{F_2}}{\text{\scriptsize $u$ fresh}}{\B, xRy,x \le z, \Left \SEQ \Right}{\B, xRy, x \le z, y \le u, zRu, \Left \SEQ \Right }$		
				}
			\end{tabular}		
		\end{minipage}

\end{center}

\bigskip

A small explanation about the rules $\rn{F1}$ and $\rn{F2}$ is below:

\begin{itemize}
	\item Rule $\rn{F1}$.

	As we mentioned, the next rule is needed to cover one of the pre-order relation conditions:

	\begin{center}
		$\vlinf{\rn{F_1}}{\text{\scriptsize $u$ fresh}}{\B, xRy, y \le z, \Left \SEQ \Right}{\B, xRy, y \le z, x \le u, uRz, \Left \SEQ \Right}$
	\end{center}

	Given $\B$ a set of relational and pre-order atoms, and given $\Left$ and $\Right$ multiset of labelled formulas, the rule $\rn{F1}$ is read from the conclusion to the premises as following: we have labels $x$, $y$ and $z$ such that $xRy$ and $y \le z$, then there exists a world $u$ such that $x \le u$ and $uRz$.

	\item Rule $\ftwo$:

	The rule that is capturing the frame condition $\rn{F2}$ is represented in the system  $\labIKp$ as the following:

	\begin{center}
	$\vlinf{\rn{F_2}}{\text{\scriptsize $u$ fresh}}{\B, xRy,x \le z, \Left \SEQ \Right}{\B, xRy, x \le z, y \le u, zRu, \Left \SEQ \Right }$	
	\end{center}

	As the rule for $\rn{F1}$, given $\B$ a set of relational and pre-order atoms, and given $\Left$ and $\Right$ multiset of labelled formulas, the rule $\ftwo$ expresses, semantically, from the conclusion to the premise that given the labels $x, y$ and $z$, if $xRy$ and $x \le z$, then there exists a variable $u$ such that $y \le u$ and $zRu$.

\end{itemize}

\subsection{Logical connectives for intuitionistic propositional logic}

To continue building a complete and sound labelled system for intuitionistic modal logic, in this small section we present the rules that are needed to express the semantics of the usual logical connectives. That means, we present the rules for the conjunction and disjunction of the system $\labIKp$:

\begin{center}
	\begin{minipage}{.95\textwidth}
		\begin{tabular}{@{\!}c@{\quad}c}
			$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
			&
			$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, \labels{x}{A \AND B}}{\B, \Left \SEQ \Right, \labels{x}{A}}{\B, \Left \SEQ \Right, \labels{x}{B}}$
			\\\\
			$\vliinf{\llabrn\OR}{}{\B, \Left, \labels{x}{A \OR B} \SEQ \Right}{\B, \Left, \labels{x}{A} \SEQ \Right}{\B, \Left, \labels{x}{B} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
			\\\\
		\end{tabular}
	\end{minipage}
\end{center}

To capture the propositional constants $\top$ and $\bot$, the system $\labIKp$ introduces the next two rules:

\begin{center}
		$\vlinf{\llabrn\bot}{}{\B, \Left, \labels{x}{\BOT} \SEQ \Right}{}$
		\hspace{8mm}
		$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, \labels{x}{\TOP}}{}$
\end{center}

You can find a small explanation about the rules $\llabrn\AND$ and  $\rlabrn\OR$ below:

\begin{itemize}
	\item Rule $\llabrn\AND$:

	%The system $\labIKp$ presents the next rule for the conjunction left:

	%\begin{center}
	%	$\vlinf{\llabrn\AND}{}{\B,\Left, \labels{x}{A \AND B} \SEQ \Right}{\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ \Right}$
	%\end{center}

	Following the explanation of the rules presented before, $\B$ is a set of relational and preorder atoms and $\Left$ and $\Right$ are multiset of labelled formulas. From the conclusion to the premise, this rule shows semantically that if the formula $A \vlan B$ is satisfied in the label $x$, then the formula $A$ is satisfied  in $x$ and the formula $B$ too.

	\item Rule $\rlabrn\OR$:

	Let the rule presented in $\labIKp$ for disjunction right be as follows:

	\begin{center}
		$\vlinf{\rlabrn\OR}{}{\B, \Left \SEQ \Right, \labels{x}{A \OR B}}{\B, \Left \SEQ \Right, \labels{x}{A}, \labels{x}{B}}$
	\end{center}

	From the conclusion to the premise we have that if the formula $A \vlor B$ is satisfied in $x$, then the formula $A$ and the formula $B$ are true in $x$.

\end{itemize}

\subsection{Modal operators}

We introduce the rules that represent the semantics of modal operators $\BOX$ and $\DIA$. The system $\labIKp$ presents the next four rules (two for $\BOX$ and two for $\DIA$): 
\begin{center}
	\begin{minipage}{.95\textwidth}
		\begin{tabular}{@{\!}c@{\quad}c}
			\\\\
			$\vlinf{\llabrn\BOX}{}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A} \SEQ \Right}{\B, x \le y, yRz, \Left, \labels{x}{\BOX A}, \labels{z}{A} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\BOX}{\text{\scriptsize $y, z$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}{\B, x \le y, y \rel z, \Left \SEQ \Right, \labels{z}{A}}$
			\\\\
			$\vlinf{\llabrn\DIA}{\text{\scriptsize $y$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}$
			&
			$\vlinf{\rlabrn\DIA}{}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{\B, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}$
			\\
		\end{tabular}
	\end{minipage}
\end{center}

\bigskip

A more detailed explanation for some rules is below:

\begin{itemize}
	\item Rule $\llabrn\BOX$.
	Continue with the same reasoning, $\B$ is a set of relational and preorder atoms and $\Left$ and $\Right$ are multiset of laballed formulas. Semantically, the rule expresses, from the conclusion to the premise, that given $x, y$ and $z$ such that $x \le y$ and $yRz$ where the formula $\BOX A$ is true in $x$, then the formula $A$ is true in $z$.

	\item Rule $\rlabrn\DIA$.
	Semantically, the diamond establishes that in a model $\M$ and a world $v$ the formula $\DIA A$ is satisfied, if and only if, there exists a world $u$ such that $vRu$ and the formula $A$ is satisfied in $u$. In order to capture this definition, the rule shows that if $xRy$ and the formula $\DIA A$ is satisfied in $x$, then the formula $A$ is true in $y$.

\end{itemize}

\subsection{Intuitionistic implication}
To finish our system $\labIKp$, we have to introduce the rules that capture the intuitionistic implication:

\bigskip

\begin{center}
	\begin{minipage}{.95\textwidth}
		\begin{tabular}{@{\!}c@{\quad}c}
			\multicolumn{2}{c}{
				\hspace{9mm}
				$\vlinf{\rlabrn\IMP}{\text{\scriptsize $y$ fresh}}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}{\B, x \le y, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}$
			}
			\\
			\multicolumn{2}{c}{
				\hspace{10mm}
				$\vliinf{\llabrn\IMP}{}{\B, x \le y, \Left, \labels{x}{A} \IMP B \SEQ \Right}{\B, x \le y, \labels{x}{A \IMP B}, \Left \SEQ \Right, \labels{y}{A}}{\B, x \le y, \Left, \labels{y}{B} \SEQ \Right}$ 
			}
		\end{tabular}
	\end{minipage}
\end{center}

\bigskip

The rule for implication right is explained below:

\begin{itemize}
	\item Rule $\rlabrn\IMP$.
	Let $\B$ be a set of relational and preorder atoms, and given $\Left$ and $\Right$ multiset of labelled formulas, this rule shows, semantically, that if the formula $A \IMP B$ is satisfied in $x$, then there exists $y$ such that $x \le y$ and the formula $A$ is true in $y$, then the formula $B$ is also true in $y$.
\end{itemize}

\section{A complete and sound system for intuitionistic modal logics}
%
Using the rules that are presented in the previous section, we obtain a proof system $\labIKp$, displayed on Figure~\ref{fig:labIKp}, for intuitionistic modal logic following the formalism of labelled sequents. 
%
Most rules are similar to the ones of Simpson~\cite{Simpson}, but some rules are even more explicitly in correspondence with the semantics by using the preorder atoms. 
%
In particular, the rules introducing the $\BOX$-operator correspond to the definition \ref{model}.
%
Furthermore, our system, unlike Simpson's system, has to incorporate the two semantic conditions ($F_1$) and ($F_2$) into the deductive rules $\rn{F_1}$ and $\rn{F_2}$, and the rules $\rn{refl}$ and $\rn{trans}$ are also necessary to ensure that the preorder atoms do behave as a preorder relation on labels.

As we mentioned, we obtain a proof system $\labIKp$ which allows us to give an extension of labelled deduction to the intuitionistic world and then we prove the next theorem:

%\begin{theorem}
%	\label{thm:sound-compl}
%	A formula $A$ is provable in the calculus $\labIKp$ if and only if $A$ is valid in every bi-relational frame.
%\end{theorem}
%
%On the one hand, we prove directly that each rule from our system is sound wrt.~bi-relational structures.
%
%On the other hand, we show that $\labIKp$ is complete wrt.~Simpson's $\lab\IK$, and the theorem then follows from Theorem~\ref{thm:simpson-sound-compl}.

%Finally we can prove the following theorem ensuring soundness and (cut-free) completeness of $\labIKp$.

\begin{theorem}\label{thm:cutfree-compl}
	%	Let $\CC$ be a set of geometric frame properties as in~\eqref{eq:cla-geometric} and $\labbrn{\CC}$ be the corresponding set of rules following schema~\eqref{eq:modal-grs}.
	%
	For any formula $A$, the following are equivalent.
	%
	\begin{enumerate}
		\item\label{i} $A$ is a theorem of $\IK$ 
		%
		\item\label{ii} $A$ is provable in $\labIKp +\labrn{cut}$ with %\quad
%		
%		\todo{}
		\smash{%\fbox{
			\scalebox{.9}{$\vliinf{\labrn{cut}}{}{\B_1, \B_2, \Left \SEQ \Right}{\B_1, \Left \SEQ \Right, \labels{z}{C}}{\B_2, \Left, \labels{z}{C} \SEQ \Right}$}
		}%}
		%

                \lutz{why is that multiplicative on $\B$ and additive on the rest???}
		\item\label{iii} $A$ is provable in $\labIKp$
		%
		\item\label{iv} $A$ is valid in every birelational frames %satisfying the properties in $\CC$.
	\end{enumerate}
\end{theorem}

%We show that each rule from our system is sound.
%
%
%We present a completeness proof for our system $\labIKp$ using the Simpson system. 
%
%The idea comes from knowing that the Simpson system is a Cut-free system, so this proof lets us know that our system is complete without the cut rule. 
%
%We show the proof by case analysis. 
%
%Most of the rules from Simpson system are the same as the rules in the system $\labIKp$, then we prove for the rules that are different.

The proof is a careful adaptation of standard techniques.
%
In particular, in order to prove (i) $\rightarrow$ (ii), we present a syntactic completeness proof with respect the Hilbert system which means that we prove all Hilbert axioms using the rules from our system $\labIKp$: we give a proof for all the axioms of propositional intuionistic logic, for the five variants of $\mathsf{k}$ axiom from the intuitionistic syntax and finally, we simulate the necesssitation rule and modus ponens. As we mentioned, in the course of the proof of (i) $\rightarrow$ (ii), we have to derive the five $\kax$ axioms. 
%
As an example, we display the derivation of $\kax[4]$ which also illustrates the need of having the rule corresponding to $F_1$ in the system.

\vspace*{-.9cm}
$$
\hspace*{-.5cm}
\scalebox{.9}{
$
\vlderivation{
	\vlin{\rlabrn\IMP}{\text{\scriptsize $y$ fresh}} {\SEQ \labels{x}{(\DIA A \IMP \BOX B) \SEQ \BOX (A \IMP B)}}{
		\vlin{\rlabrn\BOX}{\text{\scriptsize $z, w$ fresh}}{x \le y, \labels{y}{\DIA A \IMP \BOX B} \SEQ \labels{y}{\BOX (A \IMP B)}}{
			\vlin {\rlabrn\IMP}{\text{\scriptsize $u$ fresh}}{x \le y, y\le z, z \rel w, \labels{y}{\DIA A \IMP \BOX B} \SEQ \labels{w}{A \IMP B}}{
				\vlin {\color{red}{\rn{F_1}}}{}{x \le y, y \le z, w \le u, z \rel w, \labels{y}{\DIA A \SEQ \BOX B}, \labels{u}{A} \SEQ \labels{u}{B}}{
					\vlin {\rn{trans}}{}{x \le y, y \le z, w \le u, z \le t, z \rel w, t \rel u, \labels{y}{\DIA A \IMP \BOX B}, \labels{u}{A} \SEQ \labels{u}{B}}{
						\vliin {\llabrn\IMP}{}{x \le y, y \le z, w \le u, z \le t, y \le t, z \rel w, t \rel u, \labels{y}{\DIA A \IMP \BOX B}, \labels{u}{A} \SEQ \labels{u}{B}}{
							\vlin {\rlabrn\DIA}{}{x \le y, y \le z, w \le u, z \le t, y \le t, z \rel w, t \rel u, \labels{u}{A} \SEQ \labels{u}{B}, \labels{t}{\DIA A}}{
								\vlin {\rn{refl}}{}{x \le y, y \le z, w \le u, z \le t, y \le t, z \rel w, t \rel u, \labels{u}{A} \SEQ \labels{u}{B}, \labels{t}{\DIA A}, \labels{u}{A}}{
									\vlin {\labrn{id_g}}{}{x \le y, y \le z, w \le u, z \le t, y \le t, u \le u, z \rel w, t \rel u, \labels{u}{A} \SEQ \labels{u}{B}, \labels{t}{\DIA A}, \labels{u}{A}}{
										\vlhy {}
										}
									}
								}
							}{
%						\vlin {\rn{refl}}{}{x \le y, y \le z, w \le u, z \le t, y \le t, z \rel w, t \rel u, \labels{y}{\DIA A \IMP \BOX B}, \labels{u}{A}, \labels{t}{\BOX B} \SEQ \labels{u}{B}}{
%							\vlin {\llabrn\BOX}{}{x \le y, y \le z, w \le u, z \le t, y \le t, t \le t, z \rel w, t \rel u, \labels{y}{\DIA A \IMP \BOX B}, \labels{u}{A}, \labels{t}{\BOX B} \SEQ \labels{u}{B}}{
%								\vlin {\rn{refl}}{}{x \le y, y \le z, w \le u, z \le t, y \le t, t \le t, z \rel w, t \rel u, \labels{y}{\DIA A \IMP \BOX B}, \labels{u}{A}, \labels{t}{\BOX B}, \labels{u}{B} \SEQ \labels{u}{B}}{
%									\vlin {\labrn{id_g}}{}{x \le y, y \le z, w \le u, z \le t, y \le t, t \le t, u \le u, z \rel w, t \rel u, \labels{y}{\DIA A \IMP \BOX B}, \labels{u}{A}, \labels{t}{\BOX B}, \labels{u}{B} \SEQ \labels{u}{B}}{
										\vlhy {\qquad\vdots\qquad}
%										}
%									}
%								}
%							}
						}
					}
				}
			}
		}
	}
}$
}$$
%\end{example}

%\vspace*{-.5cm}

Note that our system offers only an atomic version of the identity rule, though the above derivation uses a general version of the identity rule $\rn{id_g}$ that applies to generic formulas. 
%
We therefore have to show that such a rule is admissible in our system.
%
As an example, we display one step of this admissibility proof that also illustrates the need for the rule $F_2$. The other cases are standard.

\vspace*{-.5cm}
%\begin{example}
$$
\scalebox{.9}{
$
\vlderivation{
	\vlin{\llabrn\DIA}{}{\B, x \le y, \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{\DIA A}}{
		\vlin{\color{red}{\rn{F_2}}}{}{\B, x \le y, x \rel z, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
			\vlin{\rlabrn\DIA}{}{\B, x \le y, x \rel z, z \le u, y \rel u, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}}{
				\vlin{\labrn{id_g}}{}{\B, x \le y, x \rel z, z \le u, y \rel u, \Left, \labels{z}{A} \SEQ \Right, \labels{y}{\DIA A}, \labels{u}{A}}{
					\vlhy{}
				}
			}
		}
	}
}
$
}
$$
%\end{example}

\todo{Fix the size of the next proof}
\marianela{I will do it at the end when we have the format that we want to submit}
\begin{proof}[Proof of \ref{ii} $\rightarrow$ \ref{iii}]

	By induction on number of cuts + (rank,height) of the left-most top-most cut.

	\textbf{Commutative cases:}\label{commutative-cases}

	\begin{smallequation*}
		\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
			\vliin{\llabrn\IMP}{}{\B_1, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}}{
				\vlhtr{\DD_1}{\B_1, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}, \labels{y}{A}}
			}{
			\vlhtr{\DD_2}{\B_1, x \le y, \Left, \labels{y}{B} \SEQ \Right, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_3}{\B_2, \Left, \labels{x}{A \IMP B}, \labels{z}{C} \SEQ \Right}
}
\reducesto
\end{smallequation*}

\begin{smallequation*}\hspace*{-2em}
	\vlderiibase{\llabrn\IMP}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}{
			\vlhtr{\DD_1}{\B_1, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}, \labels{y}{A}}
		}{
		\vlhtr{\DD_3^{\rn w}}{\B_2, \Left, \labels{x}{A \IMP B}, \labels{z}{C} \SEQ \Right, \labels{y}{A}}
	}
}{
\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{y}{B} \SEQ \Right}{
	\vlhtr{\DD_2}{\B_1, x \le y, \Left, \labels{y}{B} \SEQ \Right, \labels{z}{C}}
}{
\vlhtr{\DD_3[v/y]^{\invr{\llabrn\IMP}}}{\B_2, \Left, \labels{y}{B}, \labels{z}{C} \SEQ \Right}
}
}
\end{smallequation*}

We need to make sure that $y$ does not appear in $\DD_3$, before applying Lemma~\ref{lem:inv}.
%
If it does we rewrite it with a fresh variable $v$ first.

%\begin{smallequation*}
%	\reducesto
%	\vlderibase{\llabrn{cont}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
%		\vliin{\llabrn\IMP}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B} \SEQ \Right}{
%%			\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}{
%%				\vlhtr{\DD_1^{\rn w}}{\B_1, x \le y, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}, \labels{y}{A}}
%%				}{
%%				\vlhtr{\DD_3^{\rn w}}{\B_2, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B}, \labels{z}{C} \SEQ \Right}
%%				}
%				\vlhtr{\DD_4}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}
%			}{
%%			\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}{
%%				\vlhtr{\DD_2^{\rn w}}{\B_1, x \le y, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}
%%				}{
%%				\vlhtr{\DD_3^{\rn w}}{\B_2, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}
%%				}
%				\vlhtr{\DD_5}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}
%			}
%		}
%\end{smallequation*}
%
%\begin{smallalign*}&
%	\text{with } \DD_4 = \quad
%	\vlderivation{
%		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}{
%			\vlhtr{\DD_1^{\rn w}}{\B_1, x \le y, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{C}, \labels{y}{A}}
%		}{
%			\vlhtr{\DD_3^{\rn w}}{\B_2, \Left, \labels{x}{A \IMP B}, \labels{x}{A \IMP B}, \labels{z}{C} \SEQ \Right, \labels{y}{A}}
%		}
%	}
%	\\&
%	\text{and } \DD_5 = \quad
%	\vlderivation{
%		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}{
%			\vlhtr{\DD_2^{\rn w}}{\B_1, x \le y, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}
%		}{
%			\vlhtr{\DD_3^{\rn w}}{\B_2, \Left, \labels{x}{A \IMP B}, \labels{y}{B} \SEQ \Right}
%		}
%	}
%\end{smallalign*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, \Left \SEQ \Right, \labels{x}{A \IMP B}}{
		\vlin{\rlabrn\IMP}{\text{\footnotesize $x'$ fresh}}{\B_1, \Left \SEQ \Right, \labels{x}{A \IMP B}, \labels{z}{C}}{
			\vlhtr{\DD_1}{\B_1, x \le x', \Left, \labels{x'}{A} \SEQ \Right, \labels{x'}{B}, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_2}{\B_2, \Left, \labels{z}{C} \SEQ \Right, \labels{x}{A \IMP B}}
}
\end{smallequation*}

\begin{smallequation*}
	\reducesto
	\vlderibase{\rlabrn\IMP}{\text{\footnotesize $x''$ fresh (also in $\DD_2$)}}{\B_1, \B_2, \Left \SEQ \Right, \labels{x}{A \IMP B}}{
		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le x'' \Left, \labels{x''}{A} \SEQ \Right, \labels{x''}{B}}{
			\vlhtr{\DD_1[x''/x']}{\B_1, x \le x'', \Left, \labels{x''}{A} \SEQ \Right, \labels{x''}{B}, \labels{z}{C}}
		}{
		\vlhtr{\DD_2^{\invr{\rlabrn\IMP}}}{\B_2, x \le x'', \Left, \labels{z}{C}, \labels{x''}{A} \SEQ \Right, \labels{x''}{B}}
	}
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A} \SEQ \Right}{
		\vlin{\llabrn\BOX}{}{\B_1, x \le u, u \rel v, \Left, \labels{x}{\BOX} \SEQ \Right, \labels{z}{C}}{
			\vlhtr{\DD_1}{\B_1, x \le u, u \rel v, \Left, \labels{x}{\BOX A}, \labels{v}{A} \SEQ \Right, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_2}{\B_2, \Left, \labels{x}{\BOX A}, \labels{z}{C} \SEQ \Right}
}
\end{smallequation*}

\begin{smallequation*}
	\reducesto
	\vlderibase{\llabrn\BOX}{}{\B_1, \B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A} \SEQ \Right}{
		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A}, \labels{v}{A} \SEQ \Right}{
			\vlhtr{\DD_1}{\B_1, x \le u, u \rel v, \Left, \labels{x}{\BOX A}, \labels{v}{A} \SEQ \Right, \labels{z}{C}}
		}{
		\vlhtr{\DD_2^{\rn w}}{\B_2, \Left, \labels{x}{\BOX A}, \labels{v}{A}, \labels{z}{C} \SEQ \Right}
	}
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, \Left \SEQ \Right, \labels{x}{\BOX A}}{
		\vlin{\rlabrn\BOX}{\text{\footnotesize $x',y'$ fresh}}{\B_1, \Left \SEQ \Right, \labels{x}{\BOX A}, \labels{z}{C}}{
			\vlhtr{\DD_1}{\B_1, x \le x', x' \rel y' \Left \SEQ \Right, \labels{y'}{A}, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_2}{\B_2, \Left, \labels{z}{C} \SEQ \Right, \labels{x}{\BOX A}}
}
\end{smallequation*}

\begin{smallequation*}
	\reducesto
	\vlderibase{\rlabrn\BOX}{\text{\footnotesize $u,v$ fresh (also in $\DD_2$)}}{\B_1, \B_2, \Left \SEQ \Right, \labels{x}{\BOX A}}{
		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left \SEQ \Right, \labels{v}{A}}{
			\vlhtr{\DD_1}{\B_1, x \le u, u \rel v, \Left \SEQ \Right, \labels{v}{A}, \labels{z}{C}}
		}{
		\vlhtr{\DD_2^{\invr{\rlabrn\BOX}}}{\B_1, x \le u, u \rel v, \Left, \labels{z}{C} \SEQ \Right, \labels{v}{A}}
	}
}
\end{smallequation*}

\begin{smallequation*}\hspace*{-2em}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, \Left, \labels{x}{\DIA A} \SEQ \Right}{
		\vlin{\llabrn\DIA}{\text{\footnotesize $y'$ fresh}}{\B_1, \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{z}{C}}{
			\vlhtr{\DD_1}{\B_1, x \rel y', \Left, \labels{y'}{A} \SEQ \Right, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_2}{\B_2, \Left, \labels{x}{\DIA A}, \labels{z}{C} \SEQ \Right}
}
\reducesto
\vlderibase{\llabrn\DIA}{\text{\footnotesize $y''$ fresh (also in $\DD_2$)}}{\B_1, \B_2, \Left, \labels{x}{\DIA A} \SEQ \Right}{
	\vliin{\labrn{cut}}{}{\B_1, \B_2, x \rel y'', \Left, \labels{y''}{A} \SEQ \Right}{
		\vlhtr{\DD_1[y''/y']}{\B_1, x \rel y'', \Left, \labels{y''}{A} \SEQ \Right, \labels{z}{C}}
	}{
	\vlhtr{\DD_2^{\invr{\llabrn\DIA}}}{\B_2, \Left, \labels{y''}{A}, \labels{z}{C} \SEQ \Right}
}
}
\end{smallequation*}

\begin{smallequation*}\hspace*{-2em}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{
		\vlin{\rlabrn\DIA}{}{\B_1, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{z}{C}}{
			\vlhtr{\DD_1}{\B_1, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}, \labels{z}{C}}
		}
	}{
	\vlhtr{\DD_2}{\B_2, \Left, \labels{z}{C} \SEQ \Right, \labels{x}{\DIA A}}
}
\reducesto
%	\end{smallequation*}
%	
%	\begin{smallequation*}
\vlderibase{\rlabrn\DIA}{}{\B_1, \B_2, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{
	\vliin{\labrn{cut}}{}{\B_1, \B_2, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}{
		\vlhtr{\DD_1}{\B_1, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}, \labels{z}{C}}
	}{
	\vlhtr{\DD_2^{\rn w}}{\B_2, \Left, \labels{z}{C} \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}
}
}
\end{smallequation*}

\textbf{Key cases:}\label{key-cases}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{a} \SEQ \Right}{
		\vlin{\labrn{id}}{}{\B_1, x \le y, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{
			\vlhy{}	
		}
	}{
	\vlhtr{\DD_2}{\B_2, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right}		
}
\reducesto
\vlderibase{\llabrn{mon}}{}{\B_1, \B_2, x \le y, \Left, \labels{x}{a} \SEQ \Right}{
	\vlhtr{\DD_2^{\rn w}}{\B_1, \B_2, x \le y, \Left, \labels{x}{a}, \labels{y}{a} \SEQ \Right}
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left \SEQ \Right, \labels{y}{a}}{
		\vlhtr{\DD_1}{\B_1, \Left \SEQ \Right, \labels{x}{a}, \labels{y}{a}}		
	}{
	\vlin{\labrn{id}}{}{\B_2, x \le y, \Left, \labels{x}{a} \SEQ \Right, \labels{y}{a}}{
		\vlhy{}	
	}
}
\reducesto
\vlderibase{\rlabrn{mon}}{}{\B_1, \B_2, x \le y, \Left \SEQ \Right, \labels{y}{a}}{
	\vlhtr{\DD_1^{\rn w}}{\B_1, \B_2, x \le y, \Left \SEQ \Right, \labels{x}{a}, \labels{y}{a}}
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, \B_3, x \le y, \Left \SEQ \Right}{
		\vlin{\rlabrn\IMP}{}{\B_1, \Left \SEQ \Right, \labels{x}{A \IMP B}}{
			\vlhtr{\DD_1}{\B_1, x \le x', \Left, \labels{x'}{A} \SEQ \Right, \labels{x'}{B}}
		}
	}{
	\vliin{\llabrn\IMP}{}{\B_2, \B_3, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
		\vlhtr{\DD_2}{\B_2, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}
	}{
	\vlhtr{\DD_3}{\B_3, x \le y, \Left, \labels{y}{B} \SEQ \Right}
}
}
\reducesto
\end{smallequation*}

\begin{smallequation*}
	\hspace*{-2em}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, \B_3, x \le y, \Left \SEQ \Right}{
		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le y, \Left \SEQ \Right, \labels{y}{A}}{
			\vlin{\rlabrn\IMP}{}{\B_1, \Left \SEQ \Right, \labels{x}{A \IMP B}, \labels{y}{A}}{
				\vlhtr{\DD_1^{\rn w}}{\B_1, x \le x', \Left, \labels{x'}{A} \SEQ \Right, \labels{x'}{B}, \labels{y}{A}}
			}
			%			\vlhtr{\DD_{1\rn w}'}{\B_1, \Left \SEQ \Right, \labels{x}{A \IMP B}, \labels{y}{A}}
		}{
		\vlhtr{\DD_{2}}{\B_2, x \le y, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{y}{A}}
	}
}{
\vliin{\labrn{cut}}{}{\B_1, \B_3, x \le y, \Left, \labels{y}{A} \SEQ \Right}{
	\vlhtr{\DD_1[y/x']}{\B_1, x \le y, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}
}{
\vlhtr{\DD_3}{\B_3, x \le y, \Left, \labels{y}{B} \SEQ \Right}
}
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left \SEQ \Right}{
		\vlin{\rlabrn\BOX}{}{\B_1, \Left \SEQ \Right, \labels{x}{\BOX A}}{
			\vlhtr{\DD_1}{\B_1, x \le x', x' \rel y', \Left \SEQ \Right, \labels{y'}{A}}	
		}
	}{
	\vlin{\llabrn\BOX}{}{\B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A} \SEQ \Right}{
		\vlhtr{\DD_2}{\B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A}, \labels{v}{A} \SEQ \Right}
	}
}
\reducesto
\end{smallequation*}

\begin{smallequation*}\hspace*{-7em}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left \SEQ \Right}{
		%		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left \SEQ \Right, \labels{v}{A}}{
		\vlhtr{\DD_1[u/x',v/y']}{\B_1, x \le u, u \rel v, \Left \SEQ \Right, \labels{v}{A}}	
		%			}{
		%			\vlin{\llabrn\BOX}{}{\B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A} \SEQ \Right, \labels{v}{A}}{
		%				\vlin{\labrn{id}}{}{\B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A}, \labels{v}{A} \SEQ \Right, \labels{v}{A}}{
		%					\vlhy{}
		%					}
		%				}
		%			}
	}{
	\vliin{\labrn{cut}}{}{\B_1, \B_2, x \le u, u \rel v, \Left, \labels{v}{A} \SEQ \Right}{
		\vlin{\rlabrn\BOX}{}{\B_1, \Left, \labels{v}{A} \SEQ \Right, \labels{x}{\BOX A}}{
			\vlhtr{\DD_1^{\rn w}}{\B_1, x \le x', x' \rel y', \Left, \labels{v}{A} \SEQ \Right, \labels{x}{\BOX A}, \labels{y'}{A}}	
		}
	}{
	\vlhtr{\DD_2}{\B_2, x \le u, u \rel v, \Left, \labels{x}{\BOX A}, \labels{v}{A} \SEQ \Right}
}
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \rel y, \Left \SEQ \Right}{
		\vlin{\rlabrn\DIA}{}{\B_1, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}}{
			\vlhtr{\DD_1}{\B_1, x \rel y, \Left \SEQ \Right, \labels{x}{\DIA A}, \labels{y}{A}}
		}
	}{
	\vlin{\llabrn\DIA}{\text{\footnotesize $y'$ is fresh}}{\B_2, \Left, \labels{x}{\DIA A} \SEQ \Right}{
		\vlhtr{\DD_2}{\B_2, x \rel y',\Left, \labels{y'}{A} \SEQ \Right}
	}	
}
\end{smallequation*}

\begin{smallequation*}
	\vlderiibase{\labrn{cut}}{}{\B_1, \B_2, x \rel y, \Left \SEQ \Right}{
		\vliin{\labrn{cut}}{}{\B_1, \B_2, x \rel y, \Left \SEQ \Right, \labels{y}{A}}{
			\vlhtr{\DD_1}{\B_1, x \rel y, \Left \SEQ \Right, \labels{y}{A}, \labels{x}{\DIA A}}
		}{
		\vlin{\llabrn\DIA}{\text{\footnotesize $y'$ is fresh}}{\B_2, \Left, \labels{x}{\DIA A} \SEQ \Right, \labels{y}{A}}{
			\vlhtr{\DD_2^{\rn w}}{\B_2, x \rel y',\Left, \labels{y'}{A} \SEQ \Right, \labels{y}{A}}
		}
	}
}{
\vlhtr{\DD_2[y/y']}{\B_2, x \rel y,\Left, \labels{y}{A} \SEQ \Right}
}
\end{smallequation*}
\end{proof}

\begin{proof} [Proof of soundness]
	We need to prove that each sequent rule of our system $\labIKp$ is sound. We want to prove that for all model $\M$ that satisfies the premise, then $\M$ satifies the conclusion. As the proof is similar for each of the rules, we prove the soundness of some particular rules:

	\begin{itemize}
		\item Rule $\llabrn\BOX$:

		We want to prove that the rule for $\llabrn\BOX$ introduced in the Figure \ref{fig:labIKp} is sound. What we want to prove is:

		\begin{center}
			\emph{For all model $\M$, if $\M \Vdash \B,\Left, x \le y, yRz, \labels{x}{\BOX A}, z \colon A \Rightarrow \Right$, then $\M \Vdash \B, \Left, x \le y, yRz, \labels{x}{\BOX A} \Rightarrow \Right$.}
		\end{center}
		\lutz{we need to define what that means $\M\Vdash$sequent}

		By way of contradiction we want to see that if $\exists$ $\M_{0}$ such that $\M_{0} \not \Vdash$ $conclusion$,then $\M_{0} \not \Vdash premise$. More precisely for the $\llabrn\BOX$ rule we want to show that if there exists a model $\M_{0}$ such that $\M_{0} \not \Vdash \B, \Left, x \le y, \break yRz, \labels{x}{\BOX A} \Rightarrow \Right$ then $\M_{0} \not \Vdash \B,\Left, x \le y, yRz, \labels{x}{\BOX A}, \labels{z}{A} \Rightarrow \Right$.

		Then assume there exists a model $\M_{0}$ such that $\M_{0} \not \Vdash \B, \Left, x \le y,\break yRz, \labels{x}{\BOX A} \Rightarrow \Right$, i.e $\M_{0} \Vdash \B$, $\M_{0} \Vdash \Left$, $\M_{0} \Vdash x \le y$, $ \M_{0} \Vdash yRz$, $\M_{0} \Vdash \labels{x}{\BOX A}$ and  $\M_{0} \not \Vdash \Right$. Therefore, in particular, from $\M_{0} \Vdash \labels{x}{\BOX A}$ we have that for all world $y, z$ such that $x \le y$ and $yRz$, then $\M_{0} \Vdash \labels{z}{A}$. Finally, we obtain that $\M_{0} \not \Vdash \B, \Left, x \le y, yRz, \labels{x}{\BOX A}, \labels{z}{A} \Rightarrow \Right$. Then, the rule $\llabrn\BOX$ is sound.\lutz{that is unclear}

		\item Rule $\rlabrn\BOX$: By way of contradiction. Assume there exists a model $\M_{0}$ such that $\M_{0} \not \Vdash$ $conclusion$, i.e. $\M_{0} \not \Vdash \B, \Left \Rightarrow \Right, \labels{x}{\BOX A}$, which is the same as $\M_{0} \Vdash \vlan (\B,\Left)$ and  $\M_{0} \not \Vdash \vlor (\Right, \labels{x}{\BOX A})$. From $\M_{0} \not \Vdash \vlor (\Right, \labels{x}{\BOX A})$, we have that $\M_{0} \not \Vdash \Right$ and $\M_{0} \not \Vdash \labels{x}{\BOX A}$. Following the semantic definition of $\BOX$, there exists worlds $y, z$ in $\M_{0}$, such that $x \le y, yRz$ where $\M_{0} \not \Vdash \labels{z}{A}$. Therefore, $\M_{0} \not \Vdash premise$, i.e. $\M_{0} \not \Vdash \B, x \le y, yRz, \Left \Rightarrow \Right, \labels{z}{A}$. Then, the rule $\rlabrn\BOX$ is sound.

		\item Rule $\rlabrn\IMP$: By way of contradiction. Assume there exists a model $\M_{0}$ such that $\M_{0} \not \Vdash \B, \Left \Rightarrow \Right, \labels{x}{A \IMP B}$, i.e., in particular, $\M_{0} \not \Vdash \labels{x}{A \IMP B}$. Therefore, there exists $y$ such that $x \le y$ and $\M_{0} \Vdash \labels{y}{A}$ but $\M_{0} \not \Vdash \labels{y}{B}$. Then, we have $\M_{0} \not \Vdash \B, x \le y, \Left, \labels{y}{A} \Rightarrow \labels{y}{B}$. Finally, the rule $\rlabrn\IMP$ is sound.
	\end{itemize}
        \lutz{the arguments are not clear}

	The other cases are similar.
\end{proof}

\begin{lemma}\label{lem:weak}\hbox{}\quad
	\begin{enumerate}
		\item 
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left \SEQ \Right, \labels{x}{\BOT}}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\bm\BOT}}{\B, \Left \SEQ \Right}}$

		\item 
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left, \labels{x}{\TOP} \SEQ \Right}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\bm\TOP}}{\B, \Left \SEQ \Right}}$

		\item 
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left \SEQ \Right}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\rn w}}{\B, x \rel y, u \le v, \Left, \labels{z}{A} \SEQ \Right, \labels{w}{B}}}$
	\end{enumerate}
\end{lemma}

\begin{proof}
	Standard.
\end{proof}

\begin{lemma}\label{lem:inv}\hbox{}\quad
	\begin{enumerate}
		\item
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\invr{\llabrn\IMP}}}{\B, \Left, \labels{y}{B} \SEQ \Right}}$
		of the same (or smaller) height, for any label $y$ that does not appear in $\DD$.

		\item
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left \SEQ \Right, \labels{x}{A \IMP B}}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\invr{\rlabrn\IMP}}}{\B, x \le y, \Left, \labels{y}{A} \SEQ \Right, \labels{y}{B}}}$
		of the same (or smaller) height, for any label $y$ that does not appear in $\DD$.

		\item 
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left \SEQ \Right, \labels{x}{\BOX A}}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\invr{\rlabrn\BOX}}}{\B, x \le u, u \rel v, \Left \SEQ \Right, \labels{v}{A}}}$
		of the same (or smaller) height, for any label $u$ and $v$ that do not appear in $\DD$.

		\item 
		If there exists a proof 
		$\vlderivation{\vlhtr{\DD}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}}$ 
		then there exists a proof 
		$\vlderivation{\vlhtr{\DD^{\invr{\llabrn\DIA}}}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}}$
		of the same (or smaller) height, for any label $y$ that does not appear in $\DD$.

	\end{enumerate}
\end{lemma}

\begin{proof}
In each case, we reason by induction on the height of $\DD$.

\begin{itemize}
\item $\llabrn\IMP$:
For a proof of height $1$, it is straightforward.
%
For example, if $\DD = \vlinf{\labrn{id}}{}{\B, u \le v, \Left, \labels{u}{a}, \labels{x}{A \IMP B} \SEQ \Right, \labels{v}{a}}{}$, 
%
then we take $\DD^{\invr{\llabrn\IMP}}$ to be $\vlinf{\labrn{id}}{}{\B, u \le v, \Left, \labels{u}{a},\labels{y}{B} \SEQ \Right, \labels{v}{a}}{}$.

For a proof $\DD$ of height greater than $1$ we have two cases, depending on whether the last rule of $\DD$ acts on $\labels{x}{A \IMP B}$ or only on some part of the context.

First let us fix a given index $y$ that does not appear in $\DD$.

If we start with a proof
\begin{smallequation*}
	\DD = 
	\vlderivation{
		\vlin{\rn{r}}{}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
			\vlhtr{\DD'}{\B', \Left', \labels{x}{A \IMP B} \SEQ \Right'}
		}	
	}
\end{smallequation*}

Then by induction hypothesis there exists a proof 
\begin{smallequation*}
	\vlderivation{\vlhtr{{\DD'}^{\invr{\llabrn\IMP}}}{\B', \Left', \labels{y}{B} \SEQ \Right'}}
\end{smallequation*}
of the same (or smaller) height as $\DD'$ (as $y$ also does not appear in $\DD'$).

Therefore, we have the proof
\begin{smallequation*}
	\DD^{\invr{\llabrn\IMP}} = 
	\vlderivation{
		\vlin{\rn{r}}{}{\B, \Left, \labels{y}{B} \SEQ \Right}{
			\vlhtr{{\DD'}^{\invr{\llabrn\IMP}}}{\B', \Left', \labels{y}{B} \SEQ \Right'}
		}
	}
\end{smallequation*}
of the same (or smaller) height as $\DD$.

If we start with a proof 
\begin{smallequation*}
	\vlderivation{
		\vliin{\llabrn\IMP}{\text{\footnotesize $x \le z$ appears in $\B$}}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right}{
			\vlhtr{\DD_1}{\B, \Left, \labels{x}{A \IMP B} \SEQ \Right, \labels{z}{A}}
		}{
		\vlhtr{\DD_2}{\B, \Left, \labels{z}{B} \SEQ \Right}
	}
	}
\end{smallequation*}
then we take $\DD^{\invr{\llabrn\IMP}}$ to be $\vlderivation{\vlhtr{\DD_2[y/z]}{\B, \Left, \labels{y}{B} \SEQ \Right}}$ (as $y$ also does not appear in $\DD_2$) and its height is smaller than the one of $\DD$.

%
\item $\llabrn\DIA$:

If we start with a proof
\begin{smallequation*}
	\DD = 
	\vlderivation{
		\vlin{\rn{r}}{}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{
			\vlhtr{\DD'}{\B', \Left', \labels{x}{\DIA A} \SEQ \Right'}
		}	
	}
\end{smallequation*}

Then by induction hypothesis there exists a proof 
\begin{smallequation*}
	\vlderivation{\vlhtr{{\DD'}^{\invr{\llabrn\DIA}}}{\B', x \rel y, \Left', \labels{y}{A} \SEQ \Right'}}
\end{smallequation*}
of the same (or smaller) height as $\DD'$.

Therefore, we have the proof
\begin{smallequation*}
	\DD^{\invr{\llabrn\DIA}} = 
	\vlderivation{
		\vlin{\rn{r}}{}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}{
			\vlhtr{{\DD'}^{\invr{\llabrn\DIA}}}{\B', x \rel y, \Left', \labels{y}{A} \SEQ \Right'}
		}
	}
\end{smallequation*}
of the same (or smaller) height as $\DD$, for any label $y$ that does not appear in $\DD$.

If we start with a proof 
\begin{smallequation*}
	\vlderivation{
		\vlin{\llabrn\DIA}{\text{\footnotesize $y'$ fresh}}{\B, \Left, \labels{x}{\DIA A} \SEQ \Right}{
			\vlhtr{\DD'}{\B, x \rel y', \Left, \labels{y'}{A} \SEQ \Right}
		}	
	}
\end{smallequation*}
then we take $\DD^{\invr{\llabrn\DIA}}$ to be $\vlderivation{\vlhtr{\DD'[y/y']}{\B, x \rel y, \Left, \labels{y}{A} \SEQ \Right}}$ for any label $y$ that does not appear in $\DD$.

\item $\rlabrn\IMP$ and $\rlabrn\BOX$ : Similar.
\end{itemize}
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Extensions}

The main goal of this section is to generate stronger logics adding new axioms to our system. We say \emph{stronger logic} to refer to the fact that we are restricting the class of frames we want to consider, imposing some restrictions on the accessibility relation.

In~\cite{Simpson}, Simpson extends his basic sequent system for $\IK$ to the geometric axiom family. 
%
For example, you can add the following rule:
$$\scalebox{.9}{$\vlinf{\boxbrn{4}}{\text{\footnotesize $u'$ fresh}}{\B, w \rel v, v \rel u, \Left \SEQ \Right}{\B, w \rel v, v \rel u, w \rel u, \Left \SEQ \Right}$}$$
to it and obtain a sound and complete system wrt.~$\IK$ plus the axiom
$\ax{4}\colon (\DIA\DIA A \IMP \DIA A) \AND (\BOX A \IMP \BOX\BOX A)$, that is, wrt.~to all frames in which $\rel$ is transitive.

In~\cite{Plotkin}, Plotkin and Stirling give a more general correspondence result than Theorem~\ref{thm:plotkin}, that is, for intuitionistic modal logic extended with a family of axioms wrt.~some classes of bi-relational frames.
%
For example, the frames that validate the axiom $\rn{4}_\rn\DIA \colon \DIA\DIA A \IMP \DIA A$ are exactly the ones satisfying the condition:
%\begin{center}
	($\diabrn{4}$) if $w \rel v$ and $v \rel u$, there exists a $u'\in W$ s.t.~$u \le u'$ and $wRu'$.
%\end{center}

Incorporating the preorder symbol into the syntax of our sequents allows us to also obtain a sound and complete proof system for the intuitionistic modal logic extended with axiom $\rn{4}_\rn\DIA$, by designing the following rule:
$$\scalebox{.9}{$\vlinf{\diabrn{4}}{\text{\footnotesize $u'$ fresh}}{\B, w \rel v, v \rel u, \Left \SEQ \Right}{\B, w \rel v, v \rel u, u \le u', w \rel u' , \Left \SEQ \Right}$}$$

Therefore, we decompose further the formalism of labelled sequents and extend the reach of labelled deduction to the logics studied in~\cite{Plotkin}.
%
These systems enjoy cut-elimination via usual arguments. Proofs of the axiom mentioned for $\DIA$ and $\BOX$ are below:

Proof of $\ax{4_{\BOX}} \colon \BOX A \IMP \BOX\BOX A$

$\vlderivation{
	\vlin{\rlabrn\IMP}{}{\labels{x}{\BOX A \IMP \BOX\BOX A}}{	
		\vliq{\rlabrn\BOX}{}{x \le w, \labels{w}{\BOX A} \SEQ \labels{w}{\BOX\BOX A}}{
			\vlin{\rn{F_1}}{}{x \le w, w \le w', \bm{w' \rel v}, \bm{v \le v'}, v' \rel u, \labels{w}{\BOX A} \SEQ \labels{u}{A}}{
				\vlin{\rn{trans}}{}{x \le w, w \le w', w' \rel v, v \le v', v' \rel u, \bm{w' \le t}, \bm{t \rel v'}, \labels{w}{\BOX A} \SEQ \labels{u}{A}}{
					\vlin{\boxbrn{4}}{}{x \le w, w \le w', w' \rel v, v \le v', \bm{v' \rel u}, w' \le t, \bm{t \rel v'}, w \le t \labels{w}{\BOX A} \SEQ \labels{u}{A}}{
						\vlin{\labrn\BOX}{}{x \le w, w \le w', w' \rel v, v \le v', v' \rel u, w' \le t, t \rel v', w \le t, \bm{t \rel u}, \labels{w}{\BOX A} \SEQ \labels{u}{A}}{
							\vlin{\labrn{id}}{}{x \le w, w \le w', w' \rel v, v \le v', v' \rel u, w' \le t, t \rel v', w \le t, t \rel u, \labels{w}{\BOX A}, \labels{u}{\A} \SEQ \labels{u}{A}}{
								\vlhy{}
							}
						}
					}
				}
			}
		}
	}
}
$

Proof of $\ax{4_{\DIA}} \colon \DIA\DIA A \IMP \DIA A$:

$\vlderivation{
	\vlin{\rlabrn\IMP}{}{\labels{x}{\DIA\DIA A \IMP \DIA A}}{
		\vliq{\llabrn\DIA}{}{x \le w, \labels{w}{\DIA\DIA A} \SEQ \labels{w}{\DIA A}}{
			\vlin{\diasym_\rn{4}}{}{x \le w, w \rel v, v \rel u, \labels{u}{A} \SEQ \labels{w}{\DIA A}}{
				\vlin{\rlabrn\DIA}{}{x \le w, w \rel v, v \rel u, u \le u', w \rel u'  \labels{u}{A} \SEQ \labels{w}{\DIA A}}{
					\vlin{\labrn{id}}{}{x \le w, w \rel v, v \rel u, u \le u', w \rel u'  \labels{u}{A} \SEQ \labels{w}{\DIA A}, \labels{u'}{A}}{
						\vlhy{}
					}
				}
			}
		}
	}
}
$

\bigskip

Now we want to obtain a sound and complete proof system for the intuitionistic modal logic extended with axiom $\agklmn$ (also known as $\mathsf{Scott-Lemmon}$ $\mathsf{axiom}$): $\DIA^{k} \BOX^{l} A \IMP \BOX^{m}\DIA^{n} A$ (Figure \ref*{fig:gklmn}).\todo{Wrong label. why??}

Writing this axiom in both classical and intuitionistic logic in first-order language, we obtain:

$\star$ \emph{Classical case} \hspace{2.2mm} $\rightsquigarrow$ \hspace{3.7mm}$\forall x,y,z ( xR^{k}y \vlan xR^{m}z \rightarrow \exists u yR^{l}u \vlan zR^{n}u)$

$\star$ \emph{Intuitionistic case} $\rightsquigarrow$  $\forall x,y,z((xR^{k}y \vlan xR^{m}z) \IMP \exists y' (y \le y' \vlan \exists u (y'R^{l}u \vlan zR^{n} u)))$\\

\begin{figure}[h]
	\begin{center}
		$
		\xymatrix{
			& u' \ar@{.>}[ddr]^{R^l} \\
			& u \ar@{.>}[u]^{\le} \\
			w \ar@{->}[ur]^{R^k}\ar@{->}[dr]_{R^m} && x \\
			& v \ar@{.>}[ur]_{R^n}
		}
		$
	\end{center}
	\label{fig:gklmn}
	\caption{$\agklmn$ axiom for intuitionistic case}
\end{figure}

Following the idea to have a sound and complete system adding the axiom $\agklmn$ to the system $\labIKp$, we introduce the next rule:

\bigskip

\begin{center}
	$\vlderivation { \vlin {\gklmn}{y', u$ fresh$}{\B, xR^{k}y, xR^{m}z, \Left \Rightarrow \Right}{\vlhy {\B, y \le y', xR^{k}y, xR^{m}z, y'R^{l}u, zR^{n}u, \Left\Rightarrow \Right}}}$
\end{center}

\bigskip

We start proving that our new system is complete with the base case of $\mathsf{k = l = m = n = 1}$:

\begin{theorem}
	The system $\labIKp +\agklmn$ is complete for $\mathsf{k = l = m = n = 1}$.
\end{theorem}

\begin{proof}
	\begin{center}
		\scalebox{0.93}{
			$\vlderivation {\vlin {\sir}
				{y$ fresh$}
				{\Rightarrow \labels{x}{\DIA \BOX A \IMP \BOX \DIA A}}
				{\vlin {\sbr}
					{z, w $ fresh $}
					{x \le y, \labels{y}{\DIA \BOX A} \Rightarrow \labels{y}{\BOX \DIA A}}
					{\vlin {\sdl}
						{u$ fresh$}
						{x \le y, y \le z, zRw, \labels{y}{\DIA \BOX A} \Rightarrow \labels{w}{\DIA A}}
						{\vlin {\ftwo}
							{t$ fresh$}
							{x \le y, y \le z, zRw, yRu, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}}
							{\vlin {\gklmn}
								{t', j$ fresh$}
								{x \le y, y \le z, u \le t, zRw, yRu, zRt, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}}
								{\vlin {\sdr}
									{}
									{x \le y, y \le z, u \le t, t \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}}
									{\vlin {\trans}
										{}
										{x \le y, y \le z, u \le t, t \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
										{\vlin {\sbl}
											{}
											{x \le y, y \le z, u \le t, t \le t', u \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
											{\vlin {\refl}
												{}
												{x \le y, y \le z, u \le t, t \le t', u \le t', zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A}, \labels{j}{A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
												{\vlin {\ids}
													{}
													{x \le y, y \le z, u \le t, t \le t', u \le t',j \le j, zRw, yRu, zRt, t'Rj, wRj, \labels{u}{\BOX A}, \labels{j}{A} \Rightarrow \labels{w}{\DIA A}, \labels{j}{A}}
													{\vlhy {}}}}}}}}}}}}$}
	\end{center}
\end{proof}

In order to have a completeness proof for the general case, we need to introduce the rules from the Lemma~\ref{lemma:admis}.

\begin{lemma}\label{lemma:admis} The following rules are admissible in $\labIKp$:
	\begin{enumerate}
		\item{$\vlderivation {\vlin {\boxlk}{}{\B, \Left, x(\le \circ $R$)^{k}y, \labels{x}{\BOX^{k} A}\Rightarrow \Right}{\vlhy {\B, \Left, x(\le \circ $R$)^{k}y, \labels{x}{\BOX^{k} A}, \labels{z}{A} \Rightarrow \Right}}}$}
		\item{$\vlderivation{\vlin {\boxk}{}{\B, \Left \Rightarrow \Right, \labels{x}{\BOX^{k} A}}{\vlhy {\B, x(\le \circ $R$)^{k}y,\Left \Rightarrow \Right, \labels{y}{A}}}}$}
		\item{$\vlderivation { \vlin {\diamk}{}{\B, \Left, \labels{x}{\DIA^{k} A} \Rightarrow \Right}{\vlhy {\B, xR^{k}y, \Left, \labels{y}{A} \Rightarrow \Right}}}$ }
		\item{$\vlderivation { \vlin {\diamrk}{}{\B, \Left, xR^{k}y \Rightarrow \Right, \labels{x}{\DIA^{k}A}}{\vlhy {\B, \Left, xR^{k}y \Rightarrow \Right, \labels{x}{\DIA^{k}A}, \labels{y}{A}}}}$}
	\end{enumerate}
	%\textbf{Probably F2g???}
\end{lemma}

Our conjecture is that it is possible to prove completeness of the system for the general case by induction in the $k$ parameter. This proof is ongoing study.

\bibliographystyle{alpha}
\bibliography{references}

\end{document}

%\begin{figure}[h]
%\begin{center}
%$\vlderivation{\vlinf{\id}{}{\B, \Left, \labels{x}{A} \SEQ \labels{x}{A} }{}}$
%\hspace{5mm}$\vlderivation{\vlinf{\botlab}{}{\B, \Left, x\colon \bot \SEQ z\colon A}{}}$
%\hspace{5mm}$\vlderivation{\vlinf{\toplab}{}{\B, \Left \SEQ x \colon \top}{}}$
%
%\vspace{2mm}
%
%$\vlinf{\andleflab}{}{\B,\Left, x \colon \vls(A.B) \SEQ z \colon C}{\B, \Left, x\colon \vls(A.B,\labels{x}{A}, \labels{x}{B} \SEQ z \colon C)}$\hspace{5mm}$\vliinf{\andriglab}{}{\B,\Left \SEQ x \colon \vls(A.B)}{\B, \Left \SEQ \labels{x}{A}}{\B, \Left \SEQ \labels{x}{B}}$
%
%\vspace{2mm}
%$\vliinf{\orleflab}{}{\B, \Left, x \colon \vls[A.B] \SEQ  \colon C}{\B, \Left, x \colon \vls[A.B], \labels{x}{A} \SEQ z \colon C}{\B, \Left, x \colon \vls[A.B], x   \colon   B \SEQ z \colon C}$
%
%\vspace{2mm}
%
%$\vlinf{\orriglabo}{}{\B, \Left \SEQ x \colon \vls[A.B]}{\B, \Left \SEQ x   \colon   A}$
%\hspace{7mm}$\vlinf{\orriglabt}{}{\B, \Left \SEQ x \colon \vls[A.B]}{\B, \Left \SEQ x \colon  B}$
%
%\vspace{2mm}
%
%$\vliinf{\illab}{}{\B, \Left, \labels{x}{A} \IMP B \SEQ z \colon C}{\B, \Left, \labels{x}{A} \IMP B \SEQ \labels{x}{A}}{\B, \Left, \labels{x}{A} \IMP B, \labels{x}{B} \SEQ z \colon C}$
%
%\vspace{2mm}
%
%$\vlinf{\irlab}{}{\B, \Left \SEQ \labels{x}{A} \IMP B}{\B, \Left, \labels{x}{A} \SEQ \labels{x}{B}}$
%
%\vspace{2mm}
%
% $\vlderivation {\vlinf{\bllab}{}{\B, x \rel y, \Left x \colon \BOX A \SEQ z \colon B}{\B, x \rel y, \Left, x \colon \BOX A, y \colon A \SEQ z \colon B}}$
%\hspace{5mm}  $\vlinf{\brlab}{$ $y$ fresh$}{\B, \Left \SEQ x \colon \BOX A}{\B, x \rel y, \Left \SEQ y \colon A}$
%
%\vspace{2mm}
%
%$\vlinf{\dllab}{$ $y$ fresh $}{\B, \Left, x \colon \DIA A \SEQ z \colon B}{\B, x \rel y, \Left, x \colon \DIA A, y \colon A \SEQ z \colon B}$
%\hspace{5mm}$\vlinf{\drlab}{}{\B, x \rel y, \Left,  \SEQ x \colon \DIA A}{\B, x \rel y, \Left \SEQ y \colon A}$
%
%\end{center}
%\caption{System labIK}
%\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{figure}%[h]
%	
%	\begin{center}
%		
%		$\vlderivation { \vlin {\ids}{}{\B, \Left, \labels{x}{A} \SEQ x\colon a}{\vlhy {}}}$ \hspace{7mm} $\vlderivation { \vlin {\sbot}{}{\B, \Left, x \colon \bot \SEQ z\colon A}{\vlhy {}}}$
%		
%		\vspace{3mm}
%		
%		$\vlderivation {\vlin {\svlef}{}{\B, \Left, x \colon \vls(A.B) \SEQ z \colon C}{\vlhy {\B, \Left, \labels{x}{A}, \labels{x}{B} \SEQ z \colon C}}}$
%		\hspace{7mm}$\vlderivation { \vliin {\svrig}{}{\B, \Left, \SEQ x \colon \vls(A.B)}{\vlhy {\B, \Left \SEQ \labels{x}{A} }}{\vlhy {\B, \Left \SEQ \labels{x}{B}}}}$
%		
%		\vspace{3mm}
%		
%		
%		$\vlderivation {\vliin {\solef}{}{\B, \Left, x \colon \vls[A.B] \SEQ z \colon C}{\vlhy {\B, \Left, \labels{x}{A} \SEQ z \colon C}}{\vlhy {\B, \Left, \labels{x}{B} \SEQ z \colon C}}}$
%		\hspace{7mm}$\vlderivation { \vlin{\sorone}{}{\B, \Left \SEQ x \colon \vls[A.B]}{\vlhy {\B, \Left \SEQ \labels{x}{A}}}}$
%		\hspace{7mm}$\vlderivation { \vlin {\sotwo}{}{\B, \Left \SEQ x \colon \vls[A.B]}{\vlhy {\B, \Left \SEQ \labels{x}{B}}}}$
%		
%		\vspace{3mm}
%		
%		$\vlderivation {\vliin{\sil}{}{\B, \Left, \labels{x}{A} \IMP B \SEQ z \colon C}{\vlhy {\B, \Left \SEQ \labels{x}{A}}}{\vlhy {\B, \Left, \labels{x}{B} \SEQ z \colon C}}}$
%		\hspace{7mm}$\vlderivation {\vlin{\sir}{}{\B,  \Left, \labels{x}{A} \SEQ \labels{x}{B}}{\vlhy {\B, \Left, \labels{x}{A} \SEQ \labels{x}{B}}}}$
%		
%		\vspace{3mm}
%		
%		$\vlderivation { \vlin {\sbl}{}{\B, x \rel y, \Left, x \colon \BOX A \SEQ z\colon B}{\vlhy {\B, x \rel y, \Left, x \colon \BOX A, y \colon A \SEQ z\colon B}}}$
%		\hspace{7mm}$\vlderivation { \vlin {\sbr}{y$ is fresh$}{\B, \Left \SEQ x \colon \BOX A}{\vlhy {\B, x \rel y, \Left \SEQ y \colon A}}}$
%		
%		\vspace{3mm}
%		
%		$\vlderivation { \vlin{\sdl}{y$ is fresh$}{\B, \Left, x \colon \DIA A \SEQ z \colon B}{\vlhy {\B, x \rel y, \Left, y \colon A \SEQ z \colon B}}}$
%		\hspace{7mm}$\vlderivation {\vlin {\sdr}{}{\B,x \rel y, \Left \SEQ x \colon \DIA A}{\vlhy {\B, x \rel y, \Left \SEQ y \colon A }}}$
%		
%	\end{center}
%	
%	\caption{System $\lab\IK$}
%	\label{fig:labIK}
%\end{figure}
%
%\begin{theorem}[Simpson~\cite{Simpson}]
%	\label{thm:simpson-sound-compl}
%	A formula $A$ is provable in the calculus $\lab\IK$ if and only if $A$ is valid in every bi-relational frame.
%\end{theorem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{figure}%[h]
%	\small
%	\centering
%
%		$\vlinf{\rn{id}}{}{\B, \Left, \labels{x}{A} \SEQ \Right, \labels{x}{A} }{}$
%		\hspace{5mm}
%		$\vlinf{\llabrn\bot}{}{\B, \Left, x\colon \bot \SEQ \Right}{}$
%		\hspace{5mm}
%		$\vlinf{\rlabrn\top}{}{\B, \Left \SEQ \Right, x \colon \top}{}$
%%		
%%		\vspace{4mm}
%		\\[1.5ex]
%%		
%		$\vlinf{\llabrn\AND}{}{\B,\Left, x \colon \vls(A.B) \SEQ \Right}{\conjlef}$
%		\hspace{7mm}
%		$\vliinf{\rlabrn\AND}{}{\B,\Left \SEQ \Right, x \colon \vls(A.B)}{\conjrig}{\conjrigh}$
%		
%		\vspace{4mm}
%		
%		$\vliinf{\solef}{}{\B, \Left, x \colon \vls[A.B] \SEQ \Right}{\B, \Left, x   \colon   A \SEQ \Right}{\B, \Left, x   \colon   B \SEQ \Right}$
%		\hspace{7mm}
%		$\vlinf{\sorig}{}{\B, \Left \SEQ \Right, x \colon \vls[A.B]}{\B, \Left \SEQ \Right, x   \colon   A, x   \colon   B}$
%		
%		\vspace{4mm}
%		
%		$\vlinf{\sir}{$ $y$ fresh$}{\B, \Left \SEQ \Right, \labels{x}{A} \IMP B}{\B, \Left, x \le y, y \colon A \SEQ \Right, y \colon B}$
%		
%		\vspace{4mm}
%		
%		$\vliinf{\sil}{}{\B, \Left, x \le y, \labels{x}{A} \IMP B \SEQ \Right}{\B, \Left, x \le y, \labels{x}{A} \IMP B \SEQ \Right, y \colon A}{\B, \Left, x \le y, y \colon B \SEQ \Right}$
%		
%		\vspace{4mm}
%		
%		$\vlderivation {\vlinf{\sbl}{}{\B, \Left, x \le y, y \rel z, x \colon \BOX A \SEQ \Right}{\B,\Left, x \le y, y \rel z, x \colon \BOX A, z \colon A \SEQ \Right}}$
%		\hspace{5mm} $\vlinf{\sbr}{$ $y, z$ fresh$}{\B, \Left \SEQ \Right, x \colon \BOX A}{\B, \Left, x \le y, y \rel z \SEQ \Right, z \colon A}$
%		
%		
%		\vspace{4mm}
%		
%		$\vlinf{\sdl}{$ $y$ fresh $}{\B, \Left, x \colon \DIA A \SEQ \Right}{\B, \Left, x \rel y, y \colon A \SEQ \Right}$
%		\hspace{5mm}$\vlinf{\sdr}{}{\B, \Left, x \rel y \SEQ \Right, x \colon \DIA A}{\B, \Left, x \rel y \SEQ \Right, x \colon \DIA A, y \colon A}$
%		
%		
%		\vspace{2mm}
%		
%		
%		\vspace{2mm}
%		
%		$\vlinf{\refl}{}{\B, \Left \SEQ \Right}{\B, x\le x, \Left \SEQ \Right}$
%		\hspace{7mm} $\vlinf{\trans}{}{\B, x \le y, y \le z, \Left \SEQ \Right}{\B, x \le y, y \le z, x \le z, \Left \SEQ \Right}$
%		
%		
%		\vspace{2mm}
%		
%		$\vlinf{\rn{F_1}}{\text{\footnotesize $u$ fresh}}{\B, xRy, y \le z, \Left \SEQ \Right}{\B, xRy, y \le z, x \le u, uRz, \Left \SEQ \Right}$
%		%
%		$\vlinf{\rn{F_2}}{\text{\footnotesize $u$ fresh}}{\B, xRy,x \le z, \Left \SEQ \Right}{\B, xRy, x \le z, y \le u, zRu, \Left \SEQ \Right }$		
%	
%	\caption{System $\labIKp$}
%	\label{fig:labHIK}
%\end{figure}