### 2019-10-08 caleidoscope/main.tex
rewriting part of cut elim

\documentclass[a4paper]{article}

\input{new-macros}

\title{A fully labelled proof system for intuitionistic modal logics}
\author{Sonia Marin \and 
	Marianela Morales 
	\and
	Lutz Stra{\ss}burger }
\date{}

\begin{document}

  \maketitle

 \begin{abstract}
   In this paper we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. As a
   consequence it can encompass a wider range of intuitionistic modal
   logics than existing labelled systems.
   %%
   We also show an internal cut elimination proof for our system.
 \end{abstract}

\input{introduction}
\input{preliminaries}
\input{system}
\input{completeness}
\input{soundness}
\input{cut-elim}
\input{extensions}
\input{conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
%\nocite{*}
\bibliographystyle{alpha}
\bibliography{references}

\end{document}