### 2020-09-01 FullyLabelled/JLC/main.tex

\documentclass[a4paper]{article}
\usepackage{authblk}
\input{new-macros}

\title{A fully labelled proof system for intuitionistic modal logics}
\author[1]{Sonia Marin} 
\author[2,3]{Marianela Morales}
\author[3,2]{Lutz Stra{\ss}burger}
\affil[1]{University College London}
\affil[2]{Laboratoire d'informatique d'\'Ecole polytechnique}
\affil[3]{Inria Saclay}

\date{}

\begin{document}

  \maketitle

 \begin{abstract}
 	\review{Reconsider de abstract. It reads very “operational” and does not clearly state the contributions of the paper.}\lutz{Not yet done. I am waiting for Sect.7}
   In this paper we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. As a
   consequence it can encompass a wider range of intuitionistic modal
   logics than existing labelled systems.
   %%
   We also show an internal cut elimination proof for our system.
 \end{abstract}

\input{introduction}
\input{preliminaries}
\input{system}
\input{completeness}
\input{soundness}
\input{cut-elim}
\input{extensions}
\input{conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
%\nocite{*}
\bibliographystyle{alpha}
\bibliography{references}

\end{document}