### 2020-09-17 FullyLabelled/JLC/review-answer.txt
draft of response

Dear Didier,

thank you very much for the reviews. We are very grateful to the
reviewers for their helpful comments, which enabled us to considerably
improve our paper.

We took all comments and remarks into account, and below we address
the reviewers comments in detail.

One of the main concerns of Reviewer 1 is that we have not cited all
the relevant literature. We thank the reviewer for pointing out the
missing references and added them accordingly.

We anow address the other comments:

>  Perhaps even more importantly, the authors of the paper under
>  review neglect an important caveat in the design of rules of
>  contraction-free labelled sequent calculi: to guarantee that
>  contraction be admissible one needs to ensure that a certain
>  closure condition be satisfied

We already said in the submitted version that in our system
contraction is derivable from the monotonicity rules, which in turn
are admissible. We made this even more clear in the new version in
Remark XXXXX.

>  In Section 7, if the rules G_klmn are added without their
>  contracted instances, the calculus will not be complete. For
>  example, in the classical case, Euclidean transitivity ∀a, b,
>  c(aRb&aRc → bRc) has the contracted instance ∀a, b(aRb → bRb) that
>  allows for the derivation of BOX(BOX A → A). This, however, can
>  hardly be derived without the rule’s contracted instance.

In fact, the contracted instances are added. In the same way as for
other rules, occurring labels can be equal. We failed to make this
explicit in the first submitted version and corrected that mistake in
Remark XXXXXXXXXX

>  On top of this, the authors’ claim that labelled calculi ”work best
>  for logics with standard Kripke semantics” is utterly wrong:
>  labelled calculi have been developed for several variants of Kripke
>  semantics, such as Lewis’s semantics for counterfactuals,
>  neighbourhood semantics, Kripke semantics of impossible worlds, and
>  so on.

If something works best for A, it does not mean that it won't work for
B. All we tried to say was that for logics with the standard Kripke
semantics there is this close relationship between the semantics and
the labels in the proof system.

>  In conclusion, the paper does not look highly original. The
>  birelational semantics is not a novelty, nor a calculus
>  internalising that semantics appears an original idea.

We never claim the birelational semantics to be a novelty (it goes
back to the 1980s). Our paper is not about semantics but about about
proof systems. And our system allows for extenstions that could not be
studied before. We made this more clear in the new version (see
Section 7).

>  Also the reason for preferring one semantics for intuitionistic
>  modal logic over another should be discussed in detail, as
>  otherwise the authors’ work would rather look a routine exercise

We are puzzled by this comment, as we just picked the standard
semantics for intuitionistic modal logic, and semantics is not a
concern of our paper. However, we added in the conclusion some
comments on constructive modal logics, which are another variant of
adding BOX and DIAMOND to intuitionistic propositional logic.

The main concern of reviewer 2 is the lacking detail in the section on extensions. We tried to fix this

TODO

"need"

abstract