### 2020-09-29 FullyLabelled/JLC/main.tex

\documentclass[a4paper]{article}
\usepackage{authblk}
\input{new-macros}

\title{A fully labelled proof system for intuitionistic modal logics}
\author[1]{Sonia Marin} 
\author[2,3]{Marianela Morales}
\author[3,2]{Lutz Stra{\ss}burger}
\affil[1]{University College London}
\affil[2]{Laboratoire d'Informatique de l'\'Ecole Polytechnique}
\affil[3]{Inria Saclay}

\date{}

\begin{document}

  \maketitle

 \begin{abstract}
   \review{Reconsider abstract. It reads very “operational” and does not clearly state the contributions of the paper.}
   \lutz{Done?}
   \sonia{Added two setences at the beginning and modified it a bit.}
   Labelled proof theory has been famously successful for modal logics by mimicking their relational semantics within deductive systems. Simpson in particular designed a framework to study a variety of intuitionistic modal logics integrating a binary relation symbol in the syntax.
   %%
   In this paper, we present a labelled sequent system for
   intuitionistic modal logics such that there is not only one, but
   two relation symbols appearing in sequents: one for the accessibility
   relation associated with the Kripke semantics for normal modal logics and
   one for the preorder relation associated with the Kripke semantics
   for intuitionistic logic.
   %%
   This puts our system in close correspondence with the standard
   birelational Kripke semantics for intuitionistic modal logics. 
   %%
   As a consequence it can be extended with arbitrary intuitionistic Scott-Lemmon axioms. 
   %%
   We show soundness and completeness, together with an internal cut elimination proof, encompassing a wider array of intuitionistic modal logics than any existing labelled systems.
 \end{abstract}

\input{introduction}
\input{preliminaries}
\input{system}
\input{completeness}
\input{soundness}
\input{cut-elim}
\input{extensions}
\input{conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
%\nocite{*}
\bibliographystyle{alpha}
\bibliography{references}

\end{document}